---
title: "DESeq2 Pkd2-KO Analyses"
author: "Lizzy Ramsey"
date: "3/30/2022"
output:
    html_document:
      toc: true
      toc_float: true
      toc_depth: 5

---


### Purpose  
This markdown uses nf-core for RNAseq alignment, STAR salmon counts. All data from publicly available mouse kidney RNAseq data comparing PKD2-KO and WT.  
__p70__: AKA 39/GSE149739/precystic/p70, induced-KO model, samples collected at p70 when tubules were dilated with no overt cyst formation. Total n = 6 samples, 3 Pkd2fl/fl;Pax8rtTA;TetO-Cre and 3 WT  
Dataset: [GSE149739](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE149739)  
Processed data filepath: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/  

__p21__: AKA 56/GSE69556/cystic/p21, kidney-specific KO model, samples collected at p21 from full-blown cystic kidneys. Total n = 7, 3 Pkhd1-Cre; Pkd2fl/fl and 4 WT   
Dataset: [GSE69556](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE69556)  
Processed data filepath: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/  

__p28__: AKA 19/GSE134719/cystic/p28, kidney-specific KO model, samples collected at p28 from full-blown cystic kidneys. Total n = 23, 11 Pkhd1-Cre; Pkd2fl/fl and 12 WT     
Dataset: [GSE134719](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE134719)  
Processed data filepath: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/  


#### Load in libraries
```{r message=FALSE, warning = FALSE}
library(DESeq2)
library(stringr)
library(apeglm)
library(pheatmap)
library(dplyr)
library(ggrepel)
library(signatureSearch)
library(ExperimentHub); library(rhdf5)
library(biomaRt)
library(viridis)
library(here)
library(EnhancedVolcano)
source(here("code", "functions.R"))
```

### DEA p70 precystic GSE149739

#### Read in data  
Filepath to counts data: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/  
Metadata: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE149739 

```{r}
cts_39 <- readRDS(here("data", "GSE149739_salmonmerged_gene_counts.rds"))

meta_39 <- read.csv(here("data", "GSE149739_metadata.txt"), row.names = 1)
```

Remove IFT88
```{r}
ift88 <- grep("Ift88", meta_39$Genotype, ignore.case = TRUE, value = TRUE)
nonift88 <- meta_39$Genotype[!(meta_39$Genotype %in% ift88)]
meta_39 <- dplyr::filter(meta_39, Genotype %in% nonift88)
```

#### DDS
```{r}
cts_39 <- cts_39@assays@data@listData[["counts"]]
dds_39 <- salmon2dds(cts_39, meta_39)

```


#### Run DESeq2
```{r}
#relevel to make wildtype the reference instead of first alphabetical factor
dds_39$Genotype <- relevel(dds_39$Genotype, ref = "Wildtype")

#run DESeq: estimates size factors, estimates dispersion, fits negative binomial GLM's and performs Wald stats
dds_39 <- DESeq(dds_39) 
```

#### Results  

```{r}
#extract results table with log2 fold changes, p values and adjusted p values for desired coefficients
#alpha 0.05 cutoff
#res_nf <- results(dds_39, alpha = 0.05, contrast = c("Genotype", "Pkd2flflPax8rtTATetOcre", "Wildtype"))

res_nf <- lfcShrink(dds_39, coef = c("Genotype_Pkd2flflPax8rtTATetOcre_vs_Wildtype"), type="apeglm")
res_nf
```

DESeq2 results
```{r}
dsq2_39 <- as.data.frame(res_nf)

```

Save full deseq results
```{r}
write.csv(dsq2_39, file = here("res", "deseq2_outputs", "deseq2_fullres_39.csv"))
saveRDS(dsq2_39, file = here("res", "deseq2_outputs", "deseq2_fullres_39.rds"))
```

#### LFC cutoffs
Using < -2 and > 2 LFC threshold for up and downregulated genes 
```{r}
musdegs_39_UP <- uplfc(dsq2_39, lfcutoff = 1.1)
musdegs_39_DOWN <- downlfc(dsq2_39, lfcutoff = -1.1)

```

Total genes above threshold before mapping from mouse ensembl to human symbol/entrez
```{r}
dim(musdegs_39_UP) #159 at lfc > 1.3 , 194 at lfc > 1.1
dim(musdegs_39_DOWN) #43 at lfc < -1.3  , 63 at lfc <   63
```

```{r }
write.csv(musdegs_39_UP, file = here("res", "deseq2_outputs", "deseq2_upgenes_res_39.csv"))

write.csv(musdegs_39_DOWN, file = here("res", "deseq2_outputs", "deseq2_downgenes_res_39.csv"))

```


Variance stabilized transformation  
vst() uses a parametric fit for the dispersion. In this case, the closed-form expression for the variance stabilizing transformation is used by the vst function. If a local fit is used (option fitType="locfit" to estimateDispersions) a numerical integration is used instead. The transformed data should be approximated variance stabilized and also includes correction for size factors or normalization factors. The transformed data is on the log2 scale for large counts.  
  
rlog() is better to use for this dataset -- better for high count variance by sample (in this case, one sample has over 2x the reads compared to the others) and for smaller datasets "if you have many samples (e.g. 100s), the rlog function might take too long, and so the vst function will be a faster choice. The rlog and VST have similar properties, but the rlog requires fitting a shrinkage term for each sample and each gene which takes time. See the DESeq2 paper for more discussion on the differences (Love, Huber, and Anders 2014)."  
```{r}
#vsd_nf <- vst(dds_39, blind = FALSE) #if blind, could minimize large count differences due to anticipated experimental variation

rld_nf <- rlog(dds_39, blind = FALSE) 
```

PCA
```{r}
plotPCA(rld_nf, intgroup=c("Genotype")) + geom_label_repel(aes(label = rld_nf@colData@rownames), box.padding   = 0.15, label.padding = 0.15, point.padding = 0.3, segment.color = 'grey50') + scale_colour_viridis_d() + theme(axis.title.x = element_text(size = 20)) + theme_minimal()
```

#### Mouse to Human Ortholog Mapping
Map to human entrez and filter by LINCS genes
```{r}
hom_degs_UP <- lincsGenes(musdegs_39_UP, "up")
 
hom_degs_DOWN <-  lincsGenes(musdegs_39_DOWN, "down")

#should be 100 or less
dim(hom_degs_UP) #84 at lfc > 1.3
dim(hom_degs_DOWN) #24 at lfc <-1.3
```


```{r}
write.csv(hom_degs_UP, file = here("res", "deseq2_outputs", "p70_signature_UP.csv"))
write.csv(hom_degs_DOWN, file = here("res", "deseq2_outputs", "p70_signature_DOWN.csv"))
```


#### Heatmap plotting

Map measured mouse ensembl genes to mouse MGI symbol genes and then filter for coding genes (remove pseudogenes, ribosomal, and predicted)
```{r}
#retrieve counts data and ENMUSG #
cts_df <- as.data.frame(assays((rld_nf))) 
#gene_id into a column
cts_df <- tibble:::rownames_to_column(cts_df, var = "ensembl_gene_id")

#biomart to pull gene symbls and descriptions
mouse <- useEnsembl("ensembl", dataset = "mmusculus_gene_ensembl", mirror = "useast")
gene_desc <- getBM(filters = "ensembl_gene_id", attributes = c("ensembl_gene_id","entrezgene_id", "mgi_symbol", "description"), values = cts_df$ensembl_gene_id, mart = mouse)

ps <- grep("pseudogene", gene_desc$description, value = TRUE)
rp <- grep("ribosom", gene_desc$description, value = TRUE)
pred <- grep("predicted", gene_desc$description, value = TRUE)
coding_genes <- dplyr::filter(gene_desc, !description %in% c(ps) & !description %in% c(rp) & !description %in% c(pred))
```

converting gene IDs for normalized (rlog) gene counts
```{r}
#Merge gene symbols into counts and gene ID
conv_cts_all <- merge(cts_df, coding_genes, by = "ensembl_gene_id")

#Filter out repeating data
conv_cts_all <- dplyr::distinct(conv_cts_all, mgi_symbol, .keep_all = TRUE)

#Convert back to rownames but with gene symbols
conv_cts <- tibble:::column_to_rownames(conv_cts_all, "mgi_symbol")

#remove ift88
conv_cts <- subset(conv_cts, select = c("SRR11680775", "SRR11680776", "SRR11680777", "SRR11680778", "SRR11680779", "SRR11680780"))
colnames(conv_cts) <- c("WT", "WT", "WT", "PKD2", "PKD2", "PKD2")
```


greatest LFC, wrangling
```{r}
genes <-  tibble:::rownames_to_column(dsq2_39, var = "ensembl_gene_id")
#Merge gene symbols into counts and gene ID
genes <- merge(genes, coding_genes, by = "ensembl_gene_id")

#Filter out repeating data
genes <- dplyr::distinct(genes, mgi_symbol, .keep_all = TRUE)

#Convert back to rownames but with gene symbols
genes <-  tibble:::column_to_rownames(genes, "mgi_symbol")

#top 20
top20lfc <- dplyr::slice_max(genes, abs(log2FoldChange), n = 20)
```


```{r}
heatmap_p70 <- pheatmap::pheatmap(conv_cts[rownames(top20lfc), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Top 20 LFC Genes Pre-Cyst P70", angle_col = "45")

```

```{r}
write.csv(genes, file = here("res/deseq2_outputs", "deseq2_p70_fullannotatedres.csv"))

ggsave(heatmap_p70, filename = here("res", "fea", "heatmap_p70.pdf"), height = 7, width = 14)

```

gene signature
```{r}
sig_39 <- rbind(hom_degs_UP, hom_degs_DOWN)
sig_39 <- merge(sig_39, conv_cts_all, by.x = "Mouse.gene.stable.ID", by.y = "ensembl_gene_id")
#Filter out repeating data
sig_39 <- dplyr::distinct(sig_39, hgnc_symbol, .keep_all = TRUE)
sig_39_cts <- tibble::column_to_rownames(sig_39, "hgnc_symbol")

sig_39_cts <- subset(sig_39_cts, select = c("SRR11680775", "SRR11680776", "SRR11680777", "SRR11680778", "SRR11680779", "SRR11680780"))
colnames(sig_39_cts) <- c("WT", "WT", "WT", "PKD2", "PKD2", "PKD2")


heatmap_p70 <- pheatmap::pheatmap(sig_39_cts[row.names(sig_39_cts), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, fontsize_row = 3, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Gene Signature for Pre-Cyst P70", angle_col = "45")

#heatmap_p70 <- pheatmap(sig_39_cts, cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = FALSE, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Gene Signature for Pre-Cyst P70", annotation_col = colnames(sig_39_cts))

```

```{r}
write.csv(sig_39, here("res/deseq2_outputs", "signature_annotation_p70.csv"))

ggsave(heatmap_p70, filename = here("res/deseq2_outputs", "heatmap_p70_signature.pdf"))

```


### DEA p21 cystic GSE69556


#### Read in data  
__Filepath to fastq's and metadata__: /data/project/lasseigne_lab/DATASET_dir/pkdDrugRepurposing/pkd2/GSE69556/SraRunTable.txt  
__Filepath to counts data__: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/GSE69556_nfcore/results/star_salmon/salmon.merged.gene_counts.rds  
```{r}
cts_56 <- readRDS(here("data", "GSE69556_salmonmerged_gene_counts.rds"))
cts_56 <- cts_56@assays@data@listData[["counts"]]
colnames(cts_56) <- str_sub(colnames(cts_56), 1, -2) #remove random period at the end of the sample names

meta_56 <- read.csv(here("data", "GSE69556_metadata.txt"), row.names = 1)
```


#### DDS
```{r}
dds_56 <- salmon2dds(cts_56, meta_56)

```

#### Run DESeq2
```{r}
#relevel to make wildtype the reference instead of first alphabetical factor
dds_56$Genotype <- relevel(dds_56$Genotype, ref = "wildtype")

#run DESeq: estimates size factors, estimates dispersion, fits negative binomial GLM's and performs Wald stats
dds_56 <- DESeq(dds_56) #can also look at setting replaceOutliers lower for dealing with outliers 
```

#### Results  

```{r}
#extract results table with log2 fold changes, p values and adjusted p values for desired coefficients
#res_56 <- results(dds_56, alpha = 0.05, contrast = c("Genotype", "Pkhd1CrePkd2FF", "wildtype"))
res_56 <- lfcShrink(dds_56, coef = c("Genotype_Pkhd1CrePkd2FF_vs_wildtype"), type="apeglm")
res_56
```

DESeq2 results as df and by LFC threshold
```{r}
dsq2_56 <- as.data.frame(res_56)

```


```{r}
write.csv(dsq2_56, file = here("res/deseq2_outputs", "deseq2_fullres_56.csv"))
saveRDS(dsq2_56, file = here("res", "deseq2_outputs", "deseq2_fullres_56.rds"))
```


LFC cutoffs
```{r}
#NAs removed  p-adjusted value cutoff 0.05 and log2fc greater than 2.0 and less than -2.0
musdegs_56_UP <- uplfc(dsq2_56, 2.0)

musdegs_56_DOWN <- downlfc(dsq2_56, -2.0)
```

Total genes above threshold before mapping from mouse ensembl to human symbol/entrez
```{r}
dim(musdegs_56_UP) #364
dim(musdegs_56_DOWN) #57
```

Map to human entrez and filter by LINCS genes
```{r}
hom_56_UP <- lincsGenes(musdegs_56_UP, "up")

hom_56_DOWN <-  lincsGenes(musdegs_56_DOWN, "down")
#should be 100 or less
dim(hom_56_UP) #100
dim(hom_56_DOWN) #28
```


```{r}
write.csv(musdegs_56_UP, file = here("res", "deseq2_outputs", "deseq2_upgenes_res_56.csv"))
write.csv(musdegs_56_DOWN, file = here("res", "deseq2_outputs", "deseq2_downgenes_res_56.csv"))

write.csv(hom_56_UP, file = here("res", "deseq2_outputs", "p21_signature_UP.csv"))
write.csv(hom_56_DOWN, file = here("res", "deseq2_outputs", "p21_signature_DOWN.csv"))

```


Variance stabilized transformation  

```{r}
rld_56 <- rlog(dds_56, blind = FALSE) 
```

#### Visualizations  

PCA
```{r}
#regularized log transformation
plotPCA(rld_56, intgroup=c("Genotype"))
#label samples
plotPCA(rld_56, intgroup=c("Genotype")) + geom_label_repel(aes(label = rld_56@colData@rownames), box.padding   = 0.15, label.padding = 0.15, point.padding = 0.3, segment.color = 'grey50') 
```

#### Heatmap plotting

converting gene IDs for normalized (rlog) gene counts
```{r}
#retrieve counts data and ENMUSG #
cts_df_56 <- as.data.frame(assays((rld_56))) 
#gene_id into a column
cts_df_56 <- tibble:::rownames_to_column(cts_df_56, var = "ensembl_gene_id")
#Merge gene symbols into counts and gene ID
conv_cts_56_all <- merge(cts_df_56, coding_genes, by = "ensembl_gene_id")
#Filter out repeating data
conv_cts_56_all <- dplyr::distinct(conv_cts_56_all, mgi_symbol, .keep_all = TRUE)
#Convert back to rownames but with gene symbols
conv_cts_56 <- tibble:::column_to_rownames(conv_cts_56_all, "mgi_symbol")
#Remove unnecessary data
conv_cts_56 <- conv_cts_56[,!(colnames(conv_cts_56) %in% c("ensembl_gene_id", "group", "group_name", "entrezgene_id", "description"))]

#Rename the count data to correspond with experimental group or control
colnames(conv_cts_56) <- c("wildtype", "wildtype", "wildtype", "wildtype", "PKD2", "PKD2", "PKD2")
```


greatest LFC, wrangling
```{r}
genes_56 <-  tibble:::rownames_to_column(dsq2_56, var = "ensembl_gene_id")
#Merge gene symbols into counts and gene ID
genes_56 <- merge(genes_56, coding_genes, by = "ensembl_gene_id")

#Filter out repeating data
genes_56 <- dplyr::distinct(genes_56, mgi_symbol, .keep_all = TRUE)

#Convert back to rownames but with gene symbols
genes_56 <-  tibble:::column_to_rownames(genes_56, "mgi_symbol")

#top 20
top20lfc_56 <- dplyr::slice_max(genes_56, abs(log2FoldChange), n = 20)
```


```{r}
heatmap_21 <- pheatmap::pheatmap(conv_cts_56[rownames(top20lfc_56), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Top 20 LFC Genes Cystic P21", angle_col = 45)

```


```{r }
write.csv(genes_56, file = here("res/deseq2_outputs", "deseq2_p21_fullannotatedres.csv"))

ggsave(heatmap_21, filename = here("res", "fea", "heatmap_p21_top20.pdf"), height = 7, width = 14)
```

gene signature
```{r}
sig_56 <- rbind(hom_56_UP, hom_56_DOWN)
sig_56 <- merge(sig_56, conv_cts_56_all, by.x = "Mouse.gene.stable.ID", by.y = "ensembl_gene_id")
#sig_39 <- tidyr::drop_na(sig_39)
#Filter out repeating data
sig_56 <- dplyr::distinct(sig_56, hgnc_symbol, .keep_all = TRUE)
sig_56_cts <- tibble::column_to_rownames(sig_56, "hgnc_symbol")
sig_56_cts <- subset(sig_56_cts, select = c("SRR2050895", "SRR2050896", "SRR2050897", "SRR2050898", "SRR2050899", "SRR2050900", "SRR2050901"))
colnames(sig_56_cts) <- c("WT", "WT", "WT", "WT", "PKD2", "PKD2", "PKD2")



heatmap_p21 <- pheatmap::pheatmap(sig_56_cts[row.names(sig_56_cts), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, fontsize_row = 3, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Gene Signature for Cystic P21", angle_col = 45)

```


```{r}
write.csv(sig_56, here("res/deseq2_outputs", "signature_annotation_p21.csv"))
ggsave(heatmap_p21, filename = here("res/deseq2_outputs", "heatmap_p21_signature.pdf"))

```



### DEA p28 cystic GSE134719


#### Read in data  
__Filepath to fastq's and metadata__: /data/project/lasseigne_lab/DATASET_dir/pkdDrugRepurposing/pkd2/GSE134719/SraRunTable.txt  
__Filepath to counts data__: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/GSE134719_nfcore/results/star_salmon/salmon.merged.gene_counts.rds  
```{r}
cts_19 <- readRDS(here("data", "GSE134719_salmonmerged_gene_counts.rds"))

meta_19 <- read.csv(here("data", "GSE134719_metadata.txt"), row.names = 1)
```

#### DDS
```{r}
cts_19 <- cts_19@assays@data@listData[["counts"]]
dds_19 <- salmon2dds(cts_19, meta_19)
str(dds_19)
```

#### Run DESeq2
```{r}
#relevel to make wildtype the reference instead of first alphabetical factor
dds_19$Genotype <- relevel(dds_19$Genotype, ref = "wildtype")

#run DESeq: estimates size factors, estimates dispersion, fits negative binomial GLM's and performs Wald stats
dds_19 <- DESeq(dds_19) #can also look at setting replaceOutliers lower for dealing with outliers 
```

#### Results  

```{r}
#extract results table with log2 fold changes, p values and adjusted p values for desired coefficients
#res_19 <- results(dds_19, alpha = 0.05, contrast = c("Genotype", "Pdk2KO", "wildtype"))
res_19 <- lfcShrink(dds_19, coef = c("Genotype_Pdk2KO_vs_wildtype"), type="apeglm")

```

DESeq2 results as df and by LFC threshold
```{r}
dsq2_19 <- as.data.frame(res_19)

```


Save full deseq2 results
```{r}
write.csv(dsq2_19, file = here("res", "deseq2_outputs", "deseq2_fullres_19.csv"))
saveRDS(dsq2_19, file = here("res", "deseq2_outputs", "deseq2_fullres_19.rds"))
```

#### LFC cutoffs
```{r}
musdegs_19_UP <- uplfc(dsq2_19, 2.0)

musdegs_19_DOWN <- downlfc(dsq2_19, -2.0)
```

Total genes above threshold before mapping from mouse ensembl to human symbol/entrez
```{r}
dim(musdegs_19_UP) #1831
dim(musdegs_19_DOWN) #1631
```


Variance stabilized transformation  

```{r}
rld_19 <- rlog(dds_19, blind = FALSE) 
```


#### Visualizations  

PCA
```{r}
#regularized log transformation
#plotPCA(rld_19, intgroup=c("Genotype"))
#label samples
plotPCA(rld_19, intgroup=c("Genotype")) + geom_label_repel(aes(label = rld_19@colData@rownames), box.padding   = 0.15, label.padding = 0.15, point.padding = 0.3, segment.color = 'grey50') 
```


#### Mouse to Human Ortholog Mapping
Map to human entrez and filter by LINCS genes
```{r}
hom_19_UP <- lincsGenes(musdegs_19_UP, "up")

hom_19_DOWN <-  lincsGenes(musdegs_19_DOWN, "down")
#should be 100 or less
dim(hom_19_UP) #100
dim(hom_19_DOWN) #100
```


```{r}
write.csv(musdegs_19_UP, file = here("res", "deseq2_outputs", "deseq2_upgenes_res_19.csv"))
write.csv(musdegs_19_DOWN, file = here("res", "deseq2_outputs", "deseq2_downgenes_res_19.csv"))
 
write.csv(hom_19_UP, file = here("res", "deseq2_outputs", "p28_signature_UP.csv"))
write.csv(hom_19_DOWN, file = here("res", "deseq2_outputs", "p28_signature_DOWN.csv"))

```


#### Heatmap plotting

converting gene IDs for normalized (rlog) gene counts
```{r}
#retrieve counts data and ENMUSG #
cts_df_19 <- as.data.frame(assays((rld_19))) 
#gene_id into a column
cts_df_19 <- tibble:::rownames_to_column(cts_df_19, var = "ensembl_gene_id")
#Merge gene symbols into counts and gene ID
conv_cts_19_all <- merge(cts_df_19, coding_genes, by = "ensembl_gene_id")
#Filter out repeating data
conv_cts_19_all <- dplyr::distinct(conv_cts_19_all, mgi_symbol, .keep_all = TRUE)
#Convert back to rownames but with gene symbols
conv_cts_19 <- tibble:::column_to_rownames(conv_cts_19_all, "mgi_symbol")
#Remove unnecessary data
conv_cts_19 <- conv_cts_19[,!(colnames(conv_cts_19) %in% c("ensembl_gene_id", "group", "group_name", "entrezgene_id", "description"))]

#Rename the count data to correspond with experimental group or control
colnames(conv_cts_19) <- c("Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype")

```


greatest LFC, wrangling
```{r}
genes_19 <-  tibble:::rownames_to_column(dsq2_19, var = "ensembl_gene_id")
#Merge gene symbols into counts and gene ID
genes_19 <- merge(genes_19, coding_genes, by = "ensembl_gene_id")

#Filter out repeating data
genes_19 <- dplyr::distinct(genes_19, mgi_symbol, .keep_all = TRUE)

#Convert back to rownames but with gene symbols
genes_19 <-  tibble:::column_to_rownames(genes_19, "mgi_symbol")
#top 20
top20lfc_19 <- dplyr::slice_max(genes_19, abs(log2FoldChange), n = 20)
```


```{r}
heatmap_p28 <- pheatmap::pheatmap(conv_cts_19[rownames(top20lfc_19), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Top 20 LFC Genes Cystic P28", angle_col = 45)

```


```{r}
write.csv(genes_19, file = here("res/deseq2_outputs", "deseq2_p28_fullannotatedres.csv"))

ggsave(heatmap_p28, filename = here("res", "fea", "heatmap_p28.pdf"), height = 7, width = 14)
```

gene signature
```{r}
sig_wrangling <- function(lincsup, lincsdown, countconversion){
  sig <- rbind(lincsup, lincsdown)
  sig <- merge(sig, countconversion, by.x = "Mouse.gene.stable.ID", by.y = "ensembl_gene_id")

  #Filter out repeating data
  sig <- dplyr::distinct(sig_19, hgnc_symbol, .keep_all = TRUE)
  sig_cts <- tibble::column_to_rownames(sig_19, "hgnc_symbol")
  #find all sample counts
  srr_19 <- str_subset(colnames(sig_19_cts), "SRR")
  
}

sig_19 <- rbind(hom_19_UP, hom_19_DOWN)
sig_19 <- merge(sig_19, conv_cts_19_all, by.x = "Mouse.gene.stable.ID", by.y = "ensembl_gene_id")
#sig_19 <- tidyr::drop_na(sig_19)
#Filter out repeating data
sig_19 <- dplyr::distinct(sig_19, hgnc_symbol, .keep_all = TRUE)
sig_19_cts <- tibble::column_to_rownames(sig_19, "hgnc_symbol")
srr_19 <- str_subset(colnames(sig_19_cts), "SRR")
sig_19_cts <- subset(sig_19_cts, select = c(srr_19))
colnames(sig_19_cts) <- c("Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype")

heatmap_p28 <- pheatmap::pheatmap(sig_19_cts[row.names(sig_19_cts), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, fontsize_row = 2.5, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Gene Signature for Cystic P28", angle_col = 45)

```

```{r}
write.csv(sig_19, here("res/deseq2_outputs", "signature_annotation_p28.csv"))

ggsave(heatmap_p28, filename = here("res/deseq2_outputs", "heatmap_p28_signature.pdf"))

```



```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
  
