---
title: "DESeq2 GSE134719"
author: "Lizzy Ramsey"
date: "3/30/2022"
output:
    html_document:
      toc: true
      toc_float: true
      toc_depth: 5

---

```{r setup, include=FALSE}
#filepath for code and data
knitr::opts_knit$set(root.dir = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing")
```
### Purpose  
This markdown uses nf-core for RNAseq alignment and DESeq2 DDS output  
Dataset: [GSE134719](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE134719)  
Processed data filepath: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/  

#### Load in libraries
```{r message=FALSE, warning = FALSE}
library(DESeq2)
library(stringr)
library(apeglm)
library(pheatmap)
library(dplyr)
library(ggrepel)
library(signatureSearch)
library(ExperimentHub); library(rhdf5)
library(biomaRt)
```

#### Custom functions  

__salmon2dds__: Salmon merged gene counts from nf-core RNAseq pipeline (results/star_salmon/salmon.merged.gene_counts.rds) and SRA metadata as inputs, function data wrangles and runs DESeq2, resulting in a dds object with design by genotype    

```{r}
#Salmon merged gene counts to dds results
salmon2dds <- function(counts, metadata) {
  #count matrix in the sumarized experiment object
  cts <- counts@assays@data@listData[["counts"]]
  #round, salmon outputs have decimals
  cts <- round(cts)
  #colnames(cts) <- str_sub(colnames(cts), 1, -2) ##ALTERATION just for GSE69556
  #remove version numbers from transcript id's
  rownames(cts) <- gsub("\\..*", "", rownames(cts))
  
  #remove all non-alphanumeric characters from Genotype
  metadata$Genotype <- str_replace_all(metadata$Genotype, "[^[:alnum:]]", "")
  #contrasts need to be factor levels
  metadata$Genotype <- as.factor(metadata$Genotype)
#print(rownames(metadata))
#print(colnames(cts))
  dds <- DESeqDataSetFromMatrix(cts, metadata, design = ~ Genotype)
return(dds)
}
```

__uplfc__ & __downlfc__: Formats deseq2 res() output to select for upregulated and downregulated genes by log2 fold change threshold > 2 and < -2.  

```{r}
uplfc <- function(res){
  #NAs removed first
  res <- res[!is.na(res$log2FoldChange),]
  #p-adjusted value cutoff 0.05 and log2fc greater than 2.0 and less than -2.0
  upfc <- res[res$log2FoldChange > 2.0,]

  #pull just logfold change values from DESeqResults object and name the values by associated gene
  up <- upfc$log2FoldChange
  #names(up) <- rownames(upfc)
  up <- cbind(Mouse.gene.stable.ID = rownames(upfc), UpLFC = up)
  
return(as.data.frame(up))
}

downlfc <- function(res){
  #NAs removed first
  res <- res[!is.na(res$log2FoldChange),]
  #p-adjusted value cutoff 0.05 and log2fc greater than 2.0 and less than -2.0
  downfc <- res[res$log2FoldChange < -2.0,]

  #pull just logfold change values from DESeqResults object and name the values by associated gene
  down <- downfc$log2FoldChange
  #names(down) <- rownames(downfc)
  down <- cbind(Mouse.gene.stable.ID = rownames(downfc), DownLFC = down)
  
return(as.data.frame(down))
}
```

__convertMouseGeneList__: Uses Ensembl human GRCh38.p13 and Ensembl mouse C57BL_6NJ_v1 genes and to find human orthologous genes (more details [here](https://docs.google.com/document/d/1jU3EOVaZXJMzxwH9SsssHoArz8pMdcfW8MSr8hbeE4Q/edit) ), and biomart to convert to human entrez  

```{r}
convertMouseGeneList <- function(lfc){
  #read in mose to human ens annotation
  ens_annot <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/data/annot_ens_humanmouse.csv")
  #conversion table
  ens_mouse2human <- merge(lfc, ens_annot, by = "Mouse.gene.stable.ID")
  #biomart to conver human ens to human entrez
  human <- useEnsembl("ensembl", dataset = "hsapiens_gene_ensembl", mirror = "useast")
  genes <- getBM(filters = "ensembl_gene_id", attributes = c("ensembl_gene_id","entrezgene_id"), values = ens_mouse2human$Gene.stable.ID, mart = human)
  
  names(ens_mouse2human) <- c("mouse_ensembl_gene_id", "LFC", "ensembl_gene_id")
  conv_tbl <- merge(genes, ens_mouse2human, by = "ensembl_gene_id")

return(conv_tbl)
}
```

__toplincs__: Uses uplfc and downlfc outputs to then find top 100 genes that are available in LINCS (best input for signature reversion)  
```{r message=FALSE}
eh <- ExperimentHub()
lincs <- eh[["EH3226"]]; lincs_expr <- eh[["EH3227"]]

lincs_genes <- rhdf5:::h5read(lincs, "rownames", drop=TRUE)

toplincs <- function(conv_tbl, direction){
  lincs_genes <- cbind(entrezgene_id = lincs_genes)
  conv_tbl <- merge(conv_tbl, lincs_genes, by = "entrezgene_id")
  conv_tbl$LFC <- as.numeric(conv_tbl$LFC)
  if(direction == "up"){
  top100 <- slice_max(conv_tbl, n = 100, order_by = LFC)
  }
  else{
    top100 <- slice_min(conv_tbl, n = 100, order_by = LFC)
  }
return(top100)
}
```

__lincsGenes__: Wraps convertMouseGeneList to map orthologs, and toplincs to filter for LINCS-available genes 
```{r}
lincsGenes <- function(lfc, direction){
  conv_tbl <- convertMouseGeneList(lfc)
  top100 <- toplincs(conv_tbl, direction = direction)
  return(top100)
}
```

### Read in data  
__Filepath to fastq's and metadata__: /data/project/lasseigne_lab/DATASET_dir/pkdDrugRepurposing/pkd2/GSE134719/SraRunTable.txt  
__Filepath to counts data__: /data/project/lasseigne_lab/Lizzy/PKD_drugrepurposing_220324/GSE134719_nfcore/results/star_salmon/salmon.merged.gene_counts.rds  
```{r}
GSE134719_salmonmerged_gene_counts <- readRDS("~/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/data/GSE134719_salmonmerged_gene_counts.rds")

cts_19 <- GSE134719_salmonmerged_gene_counts

meta_19 <- read.csv("./data/GSE134719_metadata.txt", row.names = 1)
```

### DDS
```{r}
dds_19 <- salmon2dds(GSE134719_salmonmerged_gene_counts, meta_19)
str(dds_19)
```

### Run DESeq2
```{r}
#relevel to make wildtype the reference instead of first alphabetical factor
dds_19$Genotype <- relevel(dds_19$Genotype, ref = "wildtype")

#run DESeq: estimates size factors, estimates dispersion, fits negative binomial GLM's and performs Wald stats
dds_19 <- DESeq(dds_19) #can also look at setting replaceOutliers lower for dealing with outliers 
```

### Results  

```{r}
#extract results table with log2 fold changes, p values and adjusted p values for desired coefficients
res_19 <- results(dds_19, alpha = 0.05, contrast = c("Genotype", "Pdk2KO", "wildtype"))

```

Save full deseq2 results
```{r}
write.csv(res_19, file = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/deseq2_fullres_19.csv")
```

#### LFC cutoffs
```{r}
musdegs_19_UP <- uplfc(res_19)

musdegs_19_DOWN <- downlfc(res_19)
```

Variance stabilized transformation  

rlog() may be better to use for this dataset -- better for high count variance by sample (in this case, one sample has over 2x the reads compared to the others) and for smaller datasets "if you have many samples (e.g. 100s), the rlog function might take too long, and so the vst function will be a faster choice. The rlog and VST have similar properties, but the rlog requires fitting a shrinkage term for each sample and each gene which takes time. See the DESeq2 paper for more discussion on the differences (Love, Huber, and Anders 2014)."  
```{r}
rld_19 <- rlog(dds_19, blind = FALSE) 
```


### Visualizations  

PCA
```{r}
#regularized log transformation
#plotPCA(rld_19, intgroup=c("Genotype"))
#label samples
plotPCA(rld_19, intgroup=c("Genotype")) + geom_label_repel(aes(label = rld_19@colData@rownames), box.padding   = 0.15, label.padding = 0.15, point.padding = 0.3, segment.color = 'grey50') 
```


### Mouse to Human Ortholog Mapping
Map to human entrez and filter by LINCS genes
```{r}
hom_19_UP <- lincsGenes(musdegs_19_UP, "up")
head(hom_19_UP)
hom_19_DOWN <-  lincsGenes(musdegs_19_DOWN, "down")
head(hom_19_DOWN)
```

```{r}
write.csv(musdegs_19_UP, file = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_mus_degs_UP_220421.csv")
write.csv(musdegs_19_DOWN, file = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_mus_degs_DOWN_220421.csv")

write.csv(hom_19_UP, file = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_degs_UP_220421.csv")
write.csv(hom_19_DOWN, file = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_degs_DOWN_220421.csv")

```


__Start of Ryan's Code__

### Heatmap plotting
```{r}
#Call in pheatmap package
library(pheatmap)
#formatting for pheatmap
#Gene_name and Gene ID dataframe for merging
conversion <- as.data.frame(cts_39@elementMetadata)
#Remove .# after ENMUSG names
conversion$gene_id <- gsub("\\..*", "", conversion$gene_id)
#retrieve counts data and ENMUSG #
cts_df_19 <- as.data.frame(assays((rld_19)))
#gene_id into a column
cts_df_19 <- tibble:::rownames_to_column(cts_df_19, var = "gene_id")
#Merge gene symbols into counts and gene ID
conv_cts_19 <- merge(cts_df_19, conversion, by = "gene_id")
#Remove unnecessary data
conv_cts_19 <- conv_cts_19[,!(colnames(conv_cts_19) %in% c("gene_id", "group", "group_name"))]
#Filter out repeating data
conv_cts_19 <- dplyr::distinct(conv_cts_19, gene_name, .keep_all = TRUE)
#Convert back to rownames but with gene symbols
conv_cts_19 <- tibble:::column_to_rownames(conv_cts_19, "gene_name")
#Rename the count data to correspond with experimental group or control
colnames(conv_cts_19) <- c("Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype", "Pkd2KO", "Pkkd2KO", "Pkd2KO", "Pkd2KO", "wildtype", "wildtype", "wildtype", "wildtype")

```


greatest LFC, wrangling
```{r}
dsq2_19 <- as.data.frame(res_19)
genes_19 <-  tibble:::rownames_to_column(dsq2_19, var = "gene_id")
#Merge gene symbols into counts and gene ID
genes_19 <- merge(genes_19, conversion, by = "gene_id")
#Remove unnecessary data
genes_19 <- genes_19[,!(colnames(genes_19) %in% c("gene_id", "group", "group_name"))]
#Filter out repeating data
genes_19 <- dplyr::distinct(genes_19, gene_name, .keep_all = TRUE)
#Convert back to rownames but with gene symbols
genes_19 <-  tibble:::column_to_rownames(genes_19, "gene_name")

genes_19 <- arrange(genes_19, desc(log2FoldChange))#[1:20]
top20lfc_19 <- slice_max(genes_19, (log2FoldChange^2), n = 20)
```

```{r}
pheatmap(conv_cts_19[rownames(top20lfc_19), ], cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = TRUE, color = viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), border_color = "NA", main = "Top 20 LFC Genes Cystic P28", angle_col = 45)

```

#### Versions  
```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
  
  
  
  
