---
title: "drugtarget_analyses"
author: "Lizzy Ramsey"
date: "11/11/2022"
    html_document:
      toc: true
      toc_depth: 4
      toc_float: true
---

### Purpose  signaturesSearch analyses from pre-cystic kidney data P70 (GSE149739) compared to cystic kidney data P21 () and deseq2_GSE69556.Rmd. DEGs from these markdowns are used as input for signatureSearch. Results are annotated with Drug Repurposing Hub and onSides data.  
__Data Sets__
- '39 / Pre-Cystic P70: [GSE149739](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE149739)  
- '56 / Cystic P21: [GSE69556](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE69556)  
- '19 / Cystic P28: [GSE134719](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE134719)  

Load libraries  
```{r message=FALSE, warning = FALSE}
library(signatureSearch)
library(ggplot2)
library(ggalluvial)
#library(readr)
library(signatureSearch)
library(ExperimentHub); library(rhdf5)
library(here)
library(dplyr)
source(here("code", "functions.R"))
```

### Drug Targets in Cystic DEGs
read in DEGs from all 3 data sets (from markdown: deseq2_analyses)
```{r}
dsq2_39 <- readRDS(file = here("res", "deseq2_outputs", "deseq2_fullres_39.rds"))
dsq2_56 <- readRDS(file = here("res", "deseq2_outputs", "deseq2_fullres_56.rds"))
dsq2_19 <- readRDS(file = here("res", "deseq2_outputs", "deseq2_fullres_19.rds"))
```


Read in drug targets from FDA approved candidates from signature reversion of p70 (from markdown: sigsearch_analyses)
```{r}
fda_approved_39_targets <- read.csv(here("res", "sigsearch_outputs", "fda_drug_targets_39.csv"), row.names = 1)
```


testing new functions
```{r}
test2 <- targets_in_degs(conv_genes, dsq2_39, fda_approved_39_targets, 1.1)

#test3 <- test2 %>% dplyr::group_by(t_gn_sym) %>% stringr::str_flatten(pert, collapse = ", ")
#test3 <- stringr::str_flatten(test2$pert, collapse = ", ")

vp <- target_deg_plot(test2, 1.1, "P70 Pre-Cystic Drug Targets Found in P70 \nPre-Cystic DEGs")
vp

```


### Targets of Candidate Drugs
Targets of all FDA-approved drugs
```{r}
fda_approved_39_targets <- tidyr::separate_rows(fda_approved_39, t_gn_sym, sep = "; ")
#remove drugs with empty 'target' field
fda_approved_39_targets <- dplyr::filter(fda_approved_39_targets, t_gn_sym != "")

```

Save
```{r }
#target-separated, rhub-annotated drug results for 39
#gene_drug_targets <- rhub_launched_39_top_septarget

#write.csv(fda_approved_39_targets, file = here("res", "sigsearch_outputs", "fda_drug_targets_39.csv"))
```


Alluvial plot of drugs-targets with just FDA approved drugs
```{r}
cystic_targets <- read.csv(file = here("res/sigsearch_outputs", "drug_targets_oe_cysticdegs.csv"), row.names = 1)
#target_alluvial <- tidyr::separate_rows(cystic_overlap_drugs, t_gn_sym, sep = "; ")
#target_tbl <- cbind(pert = cystic_targets$pert, Upregulated_Target = cystic_targets$t_gn_sym.x) %>% as.data.frame()
```



plot
```{r}
#number of targets per drug
t <- table(cystic_targets$Drug_Targets) %>% as.data.frame()
t$Var1 <- as.character(t$Var1)

#cystic_targets$Drug <- as.factor(cystic_targets$Drug)
cystic_targets <- dplyr::full_join(cystic_targets, t, by= c("Drug_Targets" = "Var1"))

target_alluvial <- cystic_targets %>% mutate(Drug_Targets = forcats::fct_reorder(Drug_Targets, desc(Freq))) %>%  ggplot(aes(axis1 = Drug, axis2 = Drug_Targets)) +
  geom_alluvium(color = "black", aes(fill = Drug_Targets), show.legend=FALSE) + # , colour = trend)) + 
  geom_stratum(width = 3/12, color = "black") + #the border of the stratum
  scale_fill_viridis_d(aesthetics = c("colour", "fill")) +
  geom_text(stat = "stratum", size = 5, fontface = "bold", aes(label = after_stat(stratum))) +
  theme(axis.title.x = element_blank(), plot.title = element_text(size = 24, face = "bold", hjust = 0.5), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + ggtitle(label = "Drug Targets in P21 and P28 Cystic DEGs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank() )


target_alluvial
```




```{r}
#ggsave(here("res/sigsearch_outputs", "targets_cysticupregulated_alluvial.pdf"), target_alluvial, width = 8, height = 22)

```


Drug data for drugs whose targets are upregulated in both cystic data sets, pull repurposing hub data again, for manuscript
```{r}
cystic_overlap_drugs <- dplyr::filter(fda_approved_39, pert %in% unique(cystic_targets$Drug))

cystic_overlap_drugs <- cbind(Drug = cystic_overlap_drugs$pert, Drug_Targets = cystic_overlap_drugs$t_gn_sym, MOA = cystic_overlap_drugs$MOA, Original_Indication = cystic_overlap_drugs$Indication) %>% as.data.frame()
```

```{r}
#write.csv(cystic_overlap_drugs, file = here("res/sigsearch_outputs", "drug_summaries_oe_cystic.csv"))
```


Side effect data for drugs
```{r}
onsides_res <- dplyr::left_join(cystic_overlap_drugs, labels, by = c("Drug" = "ingredients"))

onsides_res <- dplyr::left_join(onsides_res, reactions, by = c("latest_xml_id" = "xml_id"))

liver_ade <- grep("liver", onsides_res$concept_name, ignore.case = TRUE)
liver_ade <- onsides_res[liver_ade,]

hepat_ade <- grep("hepat", onsides_res$concept_name, ignore.case = TRUE)
hepat_ade <- onsides_res[hepat_ade,]

renal_ade <- grep("renal", onsides_res$concept_name, ignore.case = TRUE)
renal_ade <- onsides_res[renal_ade,]

renal_liver_ade <- rbind(liver_ade, renal_ade)

prioritized_drugs <- dplyr::setdiff(cystic_overlap_drugs$Drug, renal_liver_ade$pert)
prioritized_drugs <- dplyr::filter(cystic_overlap_drugs, Drug %in% prioritized_drugs)
```

```{r}
#simple_tbl <- cbind(Drug = cystic_overlap_drugs$pert, Drug_Targets = cystic_overlap_drugs$t_gn_sym, MOA = cystic_overlap_drugs$MOA, Original_Indication = cystic_overlap_drugs$Indication)

#write.csv(prioritized_drugs, file = here("res/sigsearch_outputs", "drugtarget_cysticoverlap_simpletbl.csv"))
```


```{r}
#write.csv(onsides_res, file = here("res", "sigsearch_outputs", "adr_onsides_precystic_drugcandidates"))
 
```

##### Enrichment of prioritized targets, DSEA of prioritized drugs
```{r}
dsea_prioritzeddrugs <- dsea_hyperG(
  drugs = unique(cystic_targets$Drug),
  type = "GO",
  ont = "MF",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500
)

```


```{r}
dsea_prioritzeddrugs <- dsea_prioritzeddrugs@result
#write.csv(dsea_prioritzeddrugs, here("res/sigsearch_outputs", "dsea_cystictargetdrugs_pathways.csv"))
```

DSEA barplot function
```{r}
dsea_barplot <- function(dsea, titlelabel){
  dsea$Description <- stringr::str_trunc(dsea$Description, width = 75)

dseaplot <- ggplot(dplyr::slice_min(dsea, order_by = p.adjust, n = 20), aes(x=Count, y=Description, fill = p.adjust)) +
  geom_bar( stat = "identity") + 
  scale_fill_viridis_c(begin = 0.35, end = 0.95) + 
  labs(x = "Drugs Enriched for Term", y = "Functional Enrichment Terms", title = titlelabel) + 
  theme(text = element_text(size = 12), plot.title = element_text(size = 18)) + 
  theme_minimal(base_size = 14)
return(dseaplot)
}


```



```{r}
dsea_prioritzeddrugs$Description <- stringr::str_trunc(dsea_prioritzeddrugs$Description, width = 65)
dsea_prioritzeddrugs$Count <- as.numeric(dsea_prioritzeddrugs$Count)
dsea_plot <- dsea_prioritzeddrugs %>% dplyr::slice_min(order_by = p.adjust, n = 20) %>% dplyr::mutate(Description = forcats::fct_reorder(Description, (Count))) %>% ggplot( aes(x=Count, y=Description, fill = p.adjust)) +
  geom_bar( stat = "identity") + 
  scale_fill_viridis_c(begin = 0.35, end = 0.95) + 
  labs(x = "Drugs Enriched for Term", y = "Functional Enrichment Terms") + 
  theme(axis.text.x = element_text(size = 12.5, face = "bold"), axis.text.y = element_text(size = 12.5, face = "bold"), axis.title.x = element_text(size = 14, face = "bold"),  axis.title.y = element_text(size = 14, face = "bold"),plot.title = element_text(size = 16, face = "bold", hjust = 0.5)) + ggtitle(label = "Drug Target Enrichment Analysis") 
#  theme(text = element_text(size = 14, face = "bold"), plot.title = element_text(size = 18)) + 
#  theme_minimal(base_size = 16)

dsea_plot

```



```{r}

#ggsave(here("res/sigsearch_outputs", "dsea_cystictargetdrugs.png"), dsea_plot, width = 10, height = 8)

```

##### Networks: drug-target for the terms with most drugs
oxidoreductase activity, acting on CH or CH2 groups
```{r}
dtnetplot(drugs = unique(cystic_targets$Drug), set = "GO:0033695", ont = "MF", 
          desc="oxidoreductase activity, acting on CH or CH2 groups...")

```


```{r}
dtnetplot(drugs = unique(cystic_targets$Drug), set = "GO:1901338", ont = "MF", 
          desc="catecholamine binding")

```


```{r}
#dtnetplot(drugs = drugs(dsea_prioritzeddrugs), set = "GO:0099589", ont = "MF", desc="serotonin receptor activity")

```


```{r}
#dtnetplot(drugs = drugs(dsea_prioritzeddrugs), set = "GO:0004993", ont = "MF", desc="G protein-coupled serotonin receptor activity")

```


histamine receptor activity
```{r}
#dtnetplot(drugs = unique(cystic_targets$Drug), set = "GO:0004969", ont = "MF", desc="histamine receptor activity")

```

histamine receptor activity
```{r}
#dtnetplot(drugs = unique(cystic_targets$Drug), set = "GO:0050998", ont = "MF", desc="nitric-oxide synthase binding")

```


calcium-dependent protein binding
```{r}
visExport

#dtnetplot(drugs = unique(cystic_targets$Drug), set = "GO:0048306", ont = "MF", desc="calcium-dependent protein binding")

```

Signor results for prioritized 28 drug targets
```{r}
signor_res <- read.delim(here("data", "signor_prioritizedtargets.tsv"))

```


Top 16 w/ onsides
Looking at reactions for 'launched' drugs from P70/'39 results
```{r}
#onsides_top16 <- dplyr::left_join(cystic_targets, labels, by = c("Drug" = "ingredients"))
onsides_top16 <- dplyr::filter(labels, ingredients %in% unique(cystic_targets$Drug))

onsides_top16 <- dplyr::left_join(onsides_top16, reactions, by = c("latest_xml_id" = "xml_id"))

```


```{r}
#write.csv(onsides_top16, file = here("res", "sigsearch_outputs", "onsides_top16_ade.csv"))
 
```

epocrates drug prices
```{r}
avg_prices <- c(16, 14, 65, 17000, 44, 29, 28, 73, 50, 82, 19, 32, 27, 24)
#rm outlier
avg_prices <- c(16, 14, 65, 44, 29, 28, 73, 50, 82, 19, 32, 27, 24)
summary(avg_prices)
```

#### Versions  
```{r}
R.Version()
```

```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
