---
title: "sigsearch_GSE69556"
author: "Lizzy Ramsey"
date: "4/7/2022"
output:
    html_document:
      toc: true
      toc_depth: 4
      toc_float: true

---

```{r setup, include=FALSE}
#filepath for code and data
knitr::opts_knit$set(root.dir = "/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing")

```

```{r}
library(signatureSearch)
library(ggplot2)
library(ggalluvial)
```

### Purpose  
Run signatureSearch on GSE149739 data  

```{r message=FALSE, warning = FALSE}
library(readr)
library(signatureSearch)
library(ExperimentHub); library(rhdf5)

eh <- ExperimentHub()
cmap <- eh[["EH3223"]]; cmap_expr <- eh[["EH3224"]]
lincs <- eh[["EH3226"]]; lincs_expr <- eh[["EH3227"]]
h5ls(lincs)
db_path <- system.file("extdata", "sample_db.h5", package = "signatureSearch")

```

#### Custom functions  
__ssRun__: Wrapper for qsiq(), signature query, using up and down genes, and gess_lincs(), the 'lincs' signatureSearch method for gene expression signature search, sorting the results by WTCS and calculating Tau values, and then filtering for inversely related signatures and for kidney cell lines (HA1E and NKDBA) only  
```{r}
ssRun <- function(upgenes, downgenes, cells){
  q <- qSig(query = list(upset=as.character(upgenes), downset=as.character(downgenes)), gess_method="LINCS", refdb= lincs)
  gess <- gess_lincs(q, sortby="WTCS", tau=TRUE, workers=1)
  gess_res <- result(gess)
  gess_res <- dplyr::filter(gess_res, WTCS < 0)
  if(cells == "kidney"){
    gess_kidney <- gess_res %>% dplyr::filter(cell == "HA1E" | cell == "NKDBA")
    return(gess_kidney)
  }
  else{
return(gess_res)
  }
}

```

wrapper for tree_plot and ora
```{r}
#input is df with $entrez_id column
ss_fea <- function(dea){
        ora <- enrichGO(gene = dea,
                OrgDb = org.Hs.eg.db,
                keyType = "SYMBOL",
                ont = "MF", 
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.05,
                qvalueCutoff  = 0.05,
        readable      = TRUE)

return(ora)
}


#wrapper for treeplot
tplot <- function(ora, nclusters, clustmethod){
        ora_sim <- pairwise_termsim(ora)
        p1 <- treeplot(ora_sim, nCluster = nclusters)
        p2 <- treeplot(ora_sim, hclust_method = clustmethod, nCluster = nclusters)
        aplot <- aplot::plot_list(p1, p2, tag_levels='A', output = "patchwork")
return(aplot)
}

#wrapper for treeplot

#input is df with $entrez_id column

ora_tplot <- function(dea, nclusters, clustmethod){
        ora <- fea_ora(dea)
        
        aplot <- tplot(ora, nclusters, clustmethod)
return(aplot)
}

```

Group by MOA 
```{r eval=FALSE, include=FALSE}
moa_breakdown <- function(rhub_res){
  moa_list <- rhub_res$MOA
  moa_list <- stringr::str_split(moa_list, ",")
  moa_unlist <- unlist(moa_list)

  moa_df <- as.data.frame(moa_unlist)

  moa_19 <- dplyr::summarise(group_by(moa_df, MOA = moa_unlist), MOA_Freq = length(moa_unlist))

  top_moa <- dplyr::filter(moa_19, MOA_Freq > 1)
  
return(top_moa)
}

```


### GSE149739 (precystic dataset)

Load in top up and down DEGs  
```{r}
hom_degs_UP <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE149739_degs_UP_220421.csv")

hom_degs_DOWN <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE149739_degs_DOWN_220421.csv")
```

Run signatureSearch, filter for kidney cell lines  
```{r}
gess_kidneycells_39 <- ssRun(upgenes = hom_degs_UP$entrezgene_id, downgenes = hom_degs_DOWN$entrezgene_id, cells = "kidney")
```

#### Visualize Top 20
```{r}
gess_res_vis(gess_kidneycells_39, drugs = unique(gess_kidneycells_39$pert)[1:20], col = "WTCS")
```

```{r}
knitr::kable(gess_kidneycells_39, col.names = colnames(gess_kidneycells_39))

```

```{r eval=FALSE, include=FALSE}
write.csv(gess_kidneycells_39, "./res/sigsearch_outputs/gess_kidneycells_39_220421.csv")
```

## GSE69556  

Load in top up and down DEGs  
```{r}
hom_56_UP <- read_csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE69556_degs_UP_220421.csv")

hom_56_DOWN <- read_csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE69556_degs_DOWN_220421.csv")
```

Run signatureSearch, filter for kidney cell lines  
```{r}
gess_kidneycells_56 <- ssRun(upgenes = as.character(hom_56_UP$entrezgene_id), downgenes = as.character(hom_56_DOWN$entrezgene_id), cells = "kidney")
```

#### Visualize Top 20
```{r}
gess_res_vis(gess_kidneycells_56, drugs = unique(gess_kidneycells_56$pert)[1:20], col = "WTCS")
```

```{r}
knitr::kable(gess_kidneycells_56, col.names = colnames(gess_kidneycells_56))

```

Save gess and filtered gess result
```{r eval=FALSE, include=FALSE}
write.csv(gess_kidneycells_56, "./res/sigsearch_outputs/gess_kidneycells_56_220421.csv")
```

## GSE134719  
Read in data  
```{r}
hom_19_UP <- read_csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_degs_UP_220421.csv")
hom_19_DOWN <- read_csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/deseq2_outputs/GSE134719_degs_DOWN_220421.csv")
```

Run signatureSearch, filter for kidney cell lines  
```{r}
gess_kidneycells_19 <- ssRun(upgenes = as.character(hom_19_UP$entrezgene_id), downgenes = as.character(hom_19_DOWN$entrezgene_id), cells = "kidney")
```

#### Visualize Top 20
```{r}
gess_res_vis(gess_kidneycells_19, drugs = unique(gess_kidneycells_19$pert)[1:20], col = "WTCS")
```

```{r}
knitr::kable(gess_kidneycells_19, col.names = colnames(gess_kidneycells_19))
```

Save gess and filtered gess result
```{r eval=FALSE, include=FALSE}
write.csv(gess_kidneycells_19, "./res/sigsearch_outputs/gess_kidneycells_19_220421.csv")
```

## Compare datasets  
Dimensions  
```{r}
dim(gess_kidneycells_39)
dim(gess_kidneycells_56)
dim(gess_kidneycells_19)
```

Drug overlaps  
```{r}
overlap_39_56 <- dplyr::intersect(gess_kidneycells_39$pert, gess_kidneycells_56$pert) 

overlap_39_19 <- dplyr::intersect(gess_kidneycells_39$pert, gess_kidneycells_19$pert) 

overlap_56_19 <- dplyr::intersect(gess_kidneycells_56$pert, gess_kidneycells_19$pert) 

```

Number of drug overlaps  
```{r}
length(overlap_39_56)
length(overlap_39_19)
length(overlap_56_19)

length(overlap_all)
```

### SignatureSearch Plots  
```{r}
bggenes <- read.csv2(file = "./data/background_genelist.csv", row.names = 1)
#KEGG enrichment based on hypergeometric test
kegg_39 <- enrichKEGG2(gene = as.character(c(hom_degs_UP$entrezgene_id, hom_degs_DOWN$entrezgene_id)), universe = bggenes$entrezgene_id)
moa_39 <- enrichMOA(gene = as.character(c(hom_degs_UP$entrezgene_id, hom_degs_DOWN$entrezgene_id)))
```

#### TSEA  
Target Set Enrichment Analysis. Uses target protein, cellular networks and pathways affected by the top ranking drugs  

'39 TSEA 
```{r warning = FALSE}
#target set enrichment analysis TSEA, looks at target proteins for the drugs and thedrug MOA's or protein family domains
tsea_GO_39 <- tsea_dup_hyperG(drugs = gess_kidneycells_39$pert[1:100], universe = "Default", type = "GO", ont="MF", pvalueCutoff=0.05, pAdjustMethod="BH")

tsea_KEGG_39 <- tsea_dup_hyperG(drugs = gess_kidneycells_39$pert[1:50], universe = "Default", type = "KEGG", pvalueCutoff=0.05, pAdjustMethod="BH")

```


```{r warning = FALSE}
dtnetplot(drugs = drugs(tsea_KEGG_39), set = "hsa04024", ont = "MF", desc="cAMP signaling pathway")
```

```{r warning = FALSE}
dtnetplot(drugs = drugs(tsea_KEGG_39), set = "hsa04750", ont = "MF", desc="Inflammatory mediator regulation of TRP channels")
```

Save TSEA tables
```{r eval=FALSE, include=FALSE}
write.csv(tsea_GO_39@result, file = "./res/sigsearch_outputs/tsea_go_39.csv")

write.csv(tsea_KEGG_39@result, file = "./res/sigsearch_outputs/tsea_kegg_39.csv")
```

'56 TSEA 
```{r}
tsea_GO_56 <- tsea_dup_hyperG(drugs = gess_kidneycells_56$pert[1:30], universe = "Default", type = "GO", ont="MF", pvalueCutoff=0.05, pAdjustMethod="BH")

tsea_KEGG_56 <- tsea_dup_hyperG(drugs = gess_kidneycells_56$pert[1:30], universe = "Default", type = "KEGG", pvalueCutoff=0.05, pAdjustMethod="BH")

dtnetplot(drugs = drugs(tsea_GO_56), set = "GO:0004708", ont = "MF", desc="MAP kinase kinase activity")

```

```{r warning = FALSE}
dtnetplot(drugs = drugs(tsea_GO_56), set = "GO:0016712", ont = "MF", desc="oxidoreductase activity, acting on paired donors,")

dtnetplot(drugs = drugs(tsea_GO_56), set = "GO:0005262", ont = "MF", desc="calcium channel activity")
```

```{r}
dtnetplot(drugs = drugs(tsea_KEGG_56), set = "hsa04657", desc="IL-17 signaling pathway")
```

Save TSEA tables
```{r eval=FALSE, include=FALSE}
write.csv(tsea_GO_56@result, file = "./res/sigsearch_outputs/tsea_go_56.csv")

write.csv(tsea_KEGG_56@result, file = "./res/sigsearch_outputs/tsea_kegg_56.csv")
```

'19 TSEA
```{r}
tsea_GO_19 <- tsea_dup_hyperG(drugs = gess_kidneycells_19$pert[1:30], universe = "Default", type = "GO", ont="MF", pvalueCutoff=0.05, pAdjustMethod="BH")

tsea_KEGG_19 <- tsea_dup_hyperG(drugs = gess_kidneycells_19$pert[1:30], universe = "Default", type = "KEGG", pvalueCutoff=0.05, pAdjustMethod="BH")

dtnetplot(drugs = drugs(tsea_GO_19), set = "GO:0051117", ont = "MF", desc="ATPase binding")
```

```{r}
dtnetplot(drugs = drugs(tsea_GO_19), set = "GO:0051117", ont = "MF", desc="ATPase binding")

```

Save TSEA tables
```{r eval=FALSE, include=FALSE}
write.csv(tsea_GO_19@result, file = "./res/sigsearch_outputs/tsea_go_19.csv")

write.csv(tsea_KEGG_19@result, file = "./res/sigsearch_outputs/tsea_kegg_19.csv")
```

###### Plots TSEA  

Read back in  
```{r}
tsea_GO_39 <- read.csv(file = "./res/sigsearch_outputs/tsea_go_39.csv", row.names = 1)

```


TSEA gene target plot
```{r eval=FALSE, include=FALSE}
tsea_GO_39_top <- dplyr::slice_min(tsea_GO_39@result, order_by = p.adjust, n = 20)

tsea_GO_39_long <- tidyr::separate_rows(tsea_GO_39_top, itemID, sep = "/")

ggplot(data = tsea_GO_39_long,
       aes(axis1 = Description, axis2 = itemID, y = Count)) +
  geom_alluvium(aes(fill = itemID)) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Survey", "Response"),
                   expand = c(0.15, 0.05)) +
  theme_void()
```

clusterProfiler treeplots
```{r eval=FALSE, include=FALSE}
ss_drugtargets <- tidyr::separate_rows(gess_kidneycells_39, t_gn_sym, sep = "; ")
ssres_cpro <- ss_fea(unique(ss_drugtargets$t_gn_sym))

ora_sim <- pairwise_termsim(ssres_cpro)

p <- treeplot(ora_sim, hclust_method = "average", nCluster = 5, offset = rel(2), nWords = 3, label_format_tiplab = 60)
p

```

ClusterProfiler bar plot
```{r eval=FALSE, include=FALSE}
ggplot(data = slice_max(ssres_cpro@result, order_by = Count, n = 30)) +
  geom_bar(aes(x = Count, y = Description, fill = pvalue), stat = "identity") +
  scale_fill_continuous(type = "viridis") + 
  labs(x = "Number of Gene Targets", y = "Target Enrichment Terms", title = "Top Drug Candidate Target Enrichment") + 
    theme_minimal(base_size = 8)

```

#### DSEA
```{r}
dsea_39 <- dsea_hyperG(
  drugs = gess_kidneycells_39$pert[1:100],
  type = "GO",
  ont = "MF",
  pvalueCutoff = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.2,
  minGSSize = 10,
  maxGSSize = 500
)
```

```{r eval=FALSE, include=FALSE}
write.csv(dsea_39@result, file = "./res/sigsearch_outputs/dsea_39.csv")
dsea_39 <- read.csv(file = "./res/sigsearch_outputs/dsea_39.csv")
```


#### Drug Annotations  
LINCS Drug RepurposingHub, contains drug names, MOAs, original indications, and targets
```{r}
repurpHub <- read.delim("./data/Repurposing_Hub_export.txt") #6798 drugs

rHub_19 <- dplyr::inner_join(gess_kidneycells_19, repurpHub, by= c("pert" = "Name"))

rHub_56 <- dplyr::inner_join(gess_kidneycells_56, repurpHub, by= c("pert" = "Name"))

rHub_39 <- dplyr::inner_join(gess_kidneycells_39, repurpHub, by= c("pert" = "Name"))
```

P70/'39 Already-Launched Drugs
```{r}
rhub_launched_39 <- dplyr::filter(rHub_39, Phase == "Launched")

write.csv(rhub_launched_39, "./res/sigsearch_outputs/rhub_launcheddrugs_39.csv")
```

Alluvial plot of drugs-targets with just launched drugs
```{r}
rhub_launched_39_top30 <- dplyr::slice_min(rhub_launched_39, order_by = Tau, n = 30)

rhub_launched_39_top_septarget <- tidyr::separate_rows(rhub_launched_39_top30, Target, sep = ", ")
#remove drugs with empty 'target' field
rhub_launched_39_top_septarget <- dplyr::filter(rhub_launched_39_top_septarget, Target != "")
```

Save
```{r eval=FALSE, include=FALSE}
#target-separated, rhub-annotated drug results for 39
gene_drug_targets <- rhub_launched_39_top_septarget
write.csv(gene_drug_targets, file = "./res/sigsearch_outputs/gene_drug_targets_39.csv")
```


MOA's separated, string wrap labels
```{r}
rhub_launched_39_top_moasep <- tidyr::separate_rows(rhub_launched_39_top30, MOA, sep = ",")

rhub_launched_39_top_moasep$MOA <- stringr::str_wrap(rhub_launched_39_top_moasep$MOA, width = 20)

rhub_launched_39_top_moasep$Disease.Area <- stringr::str_wrap(rhub_launched_39_top_moasep$Disease.Area, width = 30)
```

plot
```{r}
moa_alluvial <- ggplot(data = rhub_launched_39_top_moasep,
       aes(axis1 = pert, axis2 = MOA)) +
  geom_alluvium(color = "black", aes(fill = WTCS)) + # , colour = trend)) + 
  geom_stratum(width = 3/12, color = "black") + #the border of the stratum
  scale_x_discrete(expand = c(.07, .07)) + 
  scale_y_discrete(expand = c(.007, .007)) + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill")) +
  geom_text(stat = "stratum", size = 3.5, fontface = "bold", aes(label = after_stat(stratum))) +
  labs(title = "MOA of Top 30 Launched Drugs") +
  theme_void() 


moa_alluvial
```

```{r eval=FALSE, include=FALSE}
ggsave("./figures/moa_alluvial_wtcs_2.pdf", moa_alluvial, width = 10, height = 15)

```


Combined repurposing hub annotated dataset drugs
```{r}
rHub_19_and_56 <-  merge(rHub_19, rHub_56, by = "pert")

rHub_all_inters <- merge(rHub_19_and_56, rHub_39,  by = "pert")
```


Save
```{r eval=FALSE, include=FALSE}
write.csv(rHub_19, file = "./res/sigsearch_outputs/ssresults_rHub_19.csv")

write.csv(rHub_56, file = "./res/sigsearch_outputs/ssresults_rHub_56.csv")

write.csv(rHub_39, file = "./res/sigsearch_outputs/ssresults_rHub_39.csv")

write.csv(rHub_19_and_56, file = "./res/sigsearch_outputs/ssresults_rHub_19_and_56.csv")

write.csv(rHub_all_inters, file = "./res/sigsearch_outputs/ssresults_rHub_all_inters.csv")

```

#### MOA Plot

Read back in
```{r}
rHub_39 <- read.csv(file = "./res/sigsearch_outputs/ssresults_rHub_39.csv", row.names = 1)

rHub_56 <- read.csv(file = "./res/sigsearch_outputs/ssresults_rHub_56.csv", row.names = 1)

rHub_19 <- read.csv(file = "./res/sigsearch_outputs/ssresults_rHub_19.csv", row.names = 1)
```

'39 / P70
```{r eval=FALSE, include=FALSE}
moa_39 <- moa_breakdown(rHub_39)

ggplot(data = moa_39) + 
  geom_bar(aes(x = MOA_Freq, y = MOA, fill = MOA), stat = "identity") +
  scale_fill_viridis_d() + 
  labs(x = "Frequency of MOA's", y = "MOA's", title = "Top MOA Categories in Drug Candidates") + 
    theme_minimal(base_size = 8)
```

Bubbleplot '39/P70
```{r eval=FALSE, include=FALSE}
ggplot(moa_39, aes(x=MOA_Freq, y=MOA, size = MOA, fill = MOA)) +
  geom_point(alpha=0.7, shape = 21) +
  scale_size(range = c(2, 10), name = "P-value") + 
  scale_fill_distiller(palette = "Purples") + 
  labs(x = "Number of Gene Targets", y = "Target Enrichment Terms") + 
  theme_minimal(base_size = 8)

```

'56 / P21
```{r eval=FALSE, include=FALSE}
moa_56 <- moa_breakdown(rHub_56)

ggplot(data = moa_56) + 
  geom_bar(aes(x = MOA_Freq, y = MOA, fill = MOA), stat = "identity") +
  scale_fill_viridis_d() + 
  labs(x = "Frequency of MOA's", y = "MOA's", title = "Top MOA Categories in Drug Candidates") + 
    theme_minimal(base_size = 8)
```

'19 / P28
```{r eval=FALSE, include=FALSE}
moa_19 <- moa_breakdown(rHub_19)

ggplot(data = moa_19) + 
  geom_bar(aes(x = MOA_Freq, y = MOA, fill = MOA), stat = "identity") +
  scale_fill_viridis_d() + 
  labs(x = "Frequency of MOA's", y = "MOA's", title = "Top MOA Categories in Drug Candidates") + 
    theme_minimal(base_size = 8)
```

##### Side effects  
Data from [onSides](https://github.com/tatonetti-lab/onsides)  
__adverse_reactions_bylabel__ - All extracted adverse reactions from the ADVERSE REACTIONS section of all labels. Each drug will have multiple labels over its lifetime (revisions, generic alternatives, etc.). This table contains the results of extracting adverse reactions from every label available for download from DailyMed. 2,764,338 rows.  
__label_map__ - Map between the xml_id (xml filename), zip_id (zip filename), and set_id (drug identifier). 45,989 rows.  

######onSides
```{r}
reactions <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/data/csv/adverse_reactions_bylabel.csv")

reactions_unique <- unique(reactions$concept_name)

warnings <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/data/csv/boxed_warnings.csv")

labels <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/data/csv/latest_labels_bydrug.csv")

```

Looking at reactions for 'launched' drugs from P70/'39 results
```{r}
onsides_39 <- dplyr::left_join(rhub_launched_39, labels, by = c("pert" = "ingredients"))

onsides_39 <- dplyr::left_join(onsides_39, reactions, by = c("latest_xml_id" = "xml_id"))

```


```{r}
write.csv(onsides_39, file = "./res/sigsearch_outputs/adr_onsides_precystic_drugcandidates")
```

#### Versions  
```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
  
