---
title: "sigsearch_analyses"
author: "Lizzy Wilk "
date: "4/7/2022"
output:
    html_document:
      toc: true
      toc_depth: 4
      toc_float: true

---


### Purpose  
This markdown follows differential expression analyses from __deseq2_analyses.Rmd__. DEGs from P70 (39) pre-cystic data are used as input for signatureSearch. Results are annotated with Drug Repurposing Hub and onSides data.  
__Data Sets__
- '39 / Pre-Cystic P70: [GSE149739](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE149739)  
- '19 / Cystic P21: [GSE134719](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE134719)  
- '56 / Cystic P28: [GSE69556](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE69556)  

Load libraries  
```{r message=FALSE, warning = FALSE}
library(signatureSearch)
library(ggplot2)
library(ggalluvial)
#library(readr)
library(signatureSearch)
library(ExperimentHub); library(rhdf5)
library(here)
library(dplyr)
library(stringr)
source(here("code", "functions.R"))
library(readr)
```

Pull LINCS data from Experiment Hub
```{r message=FALSE, warning = FALSE}
eh <- ExperimentHub()
#cmap <- eh[["EH3223"]]; cmap_expr <- eh[["EH3224"]] #cmap data locations in experimenthub
#lincs <- eh[["EH3226"]]; lincs_expr <- eh[["EH3227"]] #lincs data locations in experimenthub
#h5ls(lincs) #reading hdf5 file
#db_path <- system.file("extdata", "sample_db.h5", package = "signatureSearch")
#new lincs data
query(eh, c("signatureSearchData", "lincs2"))
lincs2_path <- eh[['EH7297']]
rhdf5::h5ls(lincs2_path) 


data("lincs_pert_info2")
#downloaded level 5 compound data from CLUE
#siginfo_beta <- fread("https://s3.amazonaws.com/macchiato.clue.io/builds/LINCS2020/siginfo_beta.txt")
#siginfo_beta <- read_delim(here("data/siginfo_beta.txt"))
#exemplar <- siginfo_beta %>% filter(pert_type=="trt_cp" & is_exemplar_sig == 1)
#new_cid <- paste(exemplar$pert_id, exemplar$cell_iname, rep("trt_cp", #length(exemplar$cmap_name)), sep="__")
#gctx2h5(here("data/level5_beta_trt_cp_n720216x12328.gctx"), cid=exemplar$sig_id, new_cid=new_cid,
#        h5file="lincs2.h5", by_ncol=5000, overwrite=TRUE)
#DBpath <- "lincs2.h5"
#sedb <- SummarizedExperiment(HDF5Array(DBpath, name="assay"))
#rownames(sedb) <- HDF5Array(DBpath, name="rownames")
#colnames(sedb) <- HDF5Array(DBpath, name="colnames")
```

### Pre-cystic (P70) Signature Reversion

Load in top up and down DEGs  
```{r}
hom_p70_UP <- read.csv(here("res", "deseq2_outputs", "p70_signature_UP.csv"))
hom_p70_DOWN <- read.csv(here("res", "deseq2_outputs", "p70_signature_DOWN.csv"))

```

Run signatureSearch, filter for kidney cell lines  
```{r}
gess_kidneycells_p70 <- ssRun(upgenes = hom_p70_UP$entrezgene_id, downgenes = hom_p70_DOWN$entrezgene_id, cells = "kidney", refdb = "lincs2") #original lincs: 100 up genes 31 down genes
```

Total results (inversely related)
```{r}
dim(gess_kidneycells_p70) #730 16
```

#### Top 30 Results by NCS
```{r}
head(gess_kidneycells_p70, 30)
```


```{r }
write.csv(gess_kidneycells_p70, file = here("res", "sigsearch_outputs", "gess_kidneycells_precystic.csv"))
 
```


### Repurposing Hub Annotations
LINCS [Drug Repurposing Hub](https://www.nature.com/articles/nm.4306), contains drug names, MOAs, original indications, and targets
```{r}
repurpHub <- read.delim(here("data", "Repurposing_Hub_export.txt"))

rHub_p70 <- dplyr::inner_join(gess_kidneycells_p70, repurpHub, by= c("pert" = "Name"))
```

P70/'39 Already-Launched Drugs
```{r}
rhub_launched_precystic <- dplyr::filter(rHub_p70, Phase == "Launched")

```

```{r}
write.csv(rhub_launched_precystic, file = here("res", "sigsearch_outputs", "rhub_launcheddrugs_precystic.csv"))
```

Alluvial plot of drugs-targets with just launched drugs
```{r}
rhub_launched_precystic_top30 <- dplyr::slice_min(rhub_launched_precystic, order_by = Tau, n = 30)

rhub_launched_precystic_top_septarget <- tidyr::separate_rows(rhub_launched_precystic_top30, Target, sep = ", ")
#remove drugs with empty 'target' field
rhub_launched_precystic_top_septarget <- dplyr::filter(rhub_launched_precystic_top_septarget, Target != "")
```

plot
```{r}
target_alluvial <- ggplot(data = rhub_launched_precystic_top_septarget,
       aes(axis1 = pert, axis2 = Target)) +
  geom_alluvium(color = "black", aes(fill = WTCS)) + # , colour = trend)) + 
  geom_stratum(width = 3/12, color = "black") + #the border of the stratum
  scale_x_discrete(expand = c(.07, .07)) + 
  scale_y_discrete(expand = c(.007, .007)) + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill")) +
  geom_text(stat = "stratum", size = 3.5, fontface = "bold", aes(label = after_stat(stratum))) +
  labs(title = "Targets of Top 20 Launched Drugs") +
  theme_void() 


target_alluvial
```


MOA's separated, string wrap labels
```{r}
rhub_launched_precystic_top_moasep <- tidyr::separate_rows(rhub_launched_precystic_top30, MOA, sep = ",")

rhub_launched_precystic_top_moasep$MOA <- stringr::str_wrap(rhub_launched_precystic_top_moasep$MOA, width = 20)

rhub_launched_precystic_top_moasep$Disease.Area <- stringr::str_wrap(rhub_launched_precystic_top_moasep$Disease.Area, width = 30)
```

plot
```{r}
moa_alluvial <- ggplot(data = rhub_launched_precystic_top_moasep,
       aes(axis1 = pert, axis2 = MOA)) +
  geom_alluvium(color = "black", aes(fill = WTCS)) + # , colour = trend)) + 
  geom_stratum(width = 3/12, color = "black") + #the border of the stratum
  scale_x_discrete(expand = c(.07, .07)) + 
  scale_y_discrete(expand = c(.007, .007)) + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill")) +
  geom_text(stat = "stratum", size = 3.5, fontface = "bold", aes(label = after_stat(stratum))) +
  labs(title = "MOA of Top 30 Launched Drugs") +
  theme_void() 


moa_alluvial
```


```{r }
ggsave(here("res", "sigsearch_outputs", "moa_alluvial_wtcs_230320.pdf"), moa_alluvial, width = 10, height = 15)

```


P70/'39 Already-Launched Drugs
```{r}
rhub_launched_p70 <- dplyr::filter(rHub_p70, Phase == "Launched")

```

Save
```{r }
write.csv(rHub_p70, file = here("res", "sigsearch_outputs", "ssresults_rHub_precystic.csv"))
```


### FDA approved drugs  
Code adapted from Jen Fisher  
Data: Drugs@FDA database was downloaded in August 2022 from https://www.fda.gov/drugs/drug-approvals-and-databases/drugsfda-data-files (Accessed date: August 2022). From the information about drugs (i.e, marketing status, application information, etc.), we created a table with information including the method of drug delivery, dosage, active ingredients, and FDA approval (FDA_products_status_220816.csv)
FDA terminology https://www.fda.gov/drugs/drug-approvals-and-databases/drugsfda-glossary-terms#:~:text=Tentative%20Approval&text=FDA%20delays%20final%20approval%20of,market%20the%20generic%20drug%20product. Drugs were filtered for unique names, regardless of form (i.e. injectible, solution, tablet, etc) and further filtered for Marketing Statuses that were not Discontinued (leaving Prescription, Over-the-counter, and Tentative Approval, due to patents)
```{r}
fda_marketing_status <- read.delim(here("data/fda_data", "MarketingStatus.txt"))
frda_product_info <- read.delim(here("data/fda_data","Products.txt"))

fda_marketing_status$ID<- paste(fda_marketing_status$ApplNo, fda_marketing_status$ProductNo, sep= "_")

frda_product_info$ID<- paste(frda_product_info$ApplNo, frda_product_info$ProductNo, sep= "_")
```

Add marketing status description to marketing status data
```{r}
marketing_lookup <- read.delim(here("data/fda_data", "MarketingStatus_Lookup.txt"))
fda_marketing_status <- merge(fda_marketing_status, marketing_lookup, by = "MarketingStatusID")
```

Combine drug info and status
```{r}
fda_data <- merge(frda_product_info, fda_marketing_status, by = "ID")
```

```{r}
write.csv(fda_data, file = here("data", "FDA_products_status_220816.csv"))
```

```{r}
fda_data <- read.csv(here("data", "FDA_products_status_220816.csv"), row.names = 1)

```

```{r}
#only unique drugs, regardless of concentration or form
fda_approved <- dplyr::filter(fda_data, MarketingStatusDescription != "Discontinued")
fda_approved <- dplyr::distinct(fda_approved, ActiveIngredient, .keep_all = TRUE)
```

```{r}
write.csv(fda_approved, file = here("data", "FDA_approved_unique_ingredients.csv"))
```

Repurposing hub 'launched' drugs
```{r}
#rhub_launched_p70 <- read.csv(here("res", "sigsearch_outputs", "rhub_launcheddrugs_p70.csv"), row.names = 1)
#fda_approved <- read.csv(here("data", "FDA_approved_unique_ingredients.csv"), row.names = 1)
```

Test for FDA approval
```{r}
fda_approved_p70 <- FDA_APPROVAL_CHECK(rhub_launched_p70$pert, fda_approved)
fda_approved_p70 <- mutate(rhub_launched_p70, fda_approval = fda_approved_p70)
fda_approved_p70 <- dplyr::filter(fda_approved_p70, fda_approval == TRUE)
#results for both kidney cell lines for some drugs
fda_approved_p70 <- dplyr::distinct(fda_approved_p70, pert, .keep_all = TRUE)
```

save FDA-approved drug candidates
```{r}
saveRDS(fda_approved_p70, file = here("res", "sigsearch_outputs", "p70_fdaapproved_drugs.rds"))
write.csv(fda_approved_p70, file = here("res", "sigsearch_outputs", "p70_fdaapproved_drugs.csv"))
```


MOA's separated, string wrap labels
```{r}
fda_approved_p70_moa <- tidyr::separate_rows(fda_approved_p70, MOA, sep = ",")

fda_approved_p70_moa_top <- fda_approved_p70_moa %>% dplyr::slice_min(order_by = NCS, n = 42) %>% mutate(pert = forcats::fct_reorder(pert, NCS))

fda_approved_p70_moa_top$MOA <- stringr::str_trunc(fda_approved_p70_moa_top$MOA, width = 38)
```


#####MOA plot
```{r}
moa_alluvial <- ggplot(fda_approved_p70_moa_top,
       aes(axis1 = pert, axis2 = MOA)) +
  geom_alluvium(color = "black", aes(fill = NCS)) + # , colour = trend)) + 
  geom_stratum(width = 5/16, color = "black") + #the border of the stratum
  scale_y_discrete(limits = ) + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill"), direction = -1) +
  geom_text(stat = "stratum", size = 3.25, fontface = "bold", aes(label = after_stat(stratum))) +
  #labs(title = "MOA of Top FDA-Approved Drugs")  + 
  theme(axis.title.x = element_blank(), plot.title = element_text(face = "bold", hjust = 0.5), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + ggtitle(label = "MOA of Top FDA-Approved Drugs") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank() )

moa_alluvial
```


```{r}
ggsave(here("res", "sigsearch_outputs", "moa_fda_alluvial.png"), moa_alluvial, width = 12, height = 7, dpi = 600)

```

Frequency of MOA's
```{r}
moa_freq <- as.data.frame(table(fda_approved_p70_moa$MOA))
other <- moa_freq$Var1[moa_freq$Freq==1]
#other <- cbind(Var1 = as.character(other), Freq = rep("other", length(other)))
multi_moa <- dplyr::filter(moa_freq, Freq > 1)
multi_moa$Var1 <- as.character(multi_moa$Var1)
moa_freqs <- rbind(multi_moa, c(Var1 = "other", Freq = length(other)))
moa_freqs$Freq <- as.numeric(moa_freqs$Freq)
#moa_freq$Var1 <- stringr::str_replace(moa_freq$Var1 %in% other == "other")

#moa_freq <- dplyr::filter(moa_freq, Freq > 1)
```


```{r}
# Stacked + percent
moa_freqs %>% mutate(MOA = forcats::fct_reorder(Var1, Freq)) %>%
ggplot(aes(fill=MOA, y=Freq, x="MOA")) + 
    geom_bar(position="fill", stat="identity") + 
  scale_fill_viridis_d(aesthetics = c("colour", "fill")) 

#visualize
moa_freq <- dplyr::filter(moa_freq, Freq > 2)
moa_freq$Var1 <- str_wrap(moa_freq$Var1, width = 25)
moa_frq <- moa_freq %>% mutate(MOA = forcats::fct_reorder(Var1, desc(Freq))) %>% ggplot(aes(x = MOA, y = Freq, fill = Freq)) +  
  geom_bar(stat = "identity") + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill")) +
  theme(axis.text.x = element_text(size = 13.5, face = "bold", angle = 50, vjust = 0.97, hjust=0.95), axis.title.x = element_blank(), plot.title = element_text(size = 14, face = "bold", hjust = 0.5)) + ggtitle(label = "Most Frequent Mechanisms of Action") + 
  theme(plot.margin = unit(c(.2,.2,.2,1.5), "cm"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  

moa_frq
```


```{r}
ggsave(filename = here("res/sigsearch_outputs", "p70_fdaapproved_moa_freq.pdf"), moa_frq, height = 5.5, width = 9.5)

```

Frequency of original disease area indications
```{r}
indication_freq <- as.data.frame(table(fda_approved_p70_moa$Disease.Area))
indication_freq <- dplyr::filter(indication_freq, Freq > 2)
```

plot
```{r}
#visualize
indication_plot <- indication_freq %>% mutate(indication = forcats::fct_reorder(Var1, desc(Freq))) %>% ggplot(aes(x = indication, y = Freq, fill = Freq)) +  
  geom_bar(stat = "identity") + 
  scale_fill_viridis_c(aesthetics = c("colour", "fill")) +
   theme(axis.text.x = element_text(size = 13.5, face = "bold", angle = 50, vjust = 0.92, hjust=0.85), axis.title.x = element_blank(), plot.title = element_text(size = 14, face = "bold", hjust = 0.5)) + ggtitle(label = "Most Frequent Disease Area Indications") + 
  theme(plot.margin = unit(c(.2,.2,.2,1), "cm"), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))


indication_plot
```

```{r}
ggsave(filename =here("res/sigsearch_outputs", "p70_fdaapproved_origindication_freq.pdf"), indication_plot, height = 5.5, width = 8)

```

#### Versions  
```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
  
