---
title: "tissue_sex_specificity"
author: "Lizzy Wilk"
date: "12/12/2022"
output: html_document
---

load libraries
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ComplexHeatmap)
library(viridis)
library(RColorBrewer)
library(ggplot2)
library(ggalluvial)
library(circlize)
```

__gtex_readin__: function to read in gtex top 500 sex-biased DEGs and add column for the tissue name  
__comb_tissues__: function to combine tissue tables and filter genes for drug targets  
__targetsintissues__:  function to data wrangle output of comb_tissues, returns a matrix of just MASH beta values (from sex bias LFC, < 0 = M, > 0 = F) for each tissue 
```{r}
gtex_readin <- function(gtex_acron, exppath = "/Users/eramsey/Downloads/PKD/kidney_drugrepurposing/data/gtex_sexspecificity_s2.xlsx", metapath = "/Users/eramsey/Downloads/PKD/kidney_drugrepurposing/data/gtex_sexspecificity_s1.xlsx"){
  gtex <- read_excel(exppath, sheet = gtex_acron, col_names = TRUE)
  #gtex <- gtex %>% dplyr::mutate(tissue = tissue) 
  meta <- read_excel(metapath, sheet = 2, col_names = TRUE)

  meta_sub <- meta %>% dplyr::filter(`Abbreviation Id` == gtex_acron)
  gtex <- gtex %>% dplyr::mutate(Tissue = meta_sub$Id) %>% dplyr::left_join( meta_sub, by = c("Tissue" = "Id"))
  
return(gtex)
}
#adrenal_gland <- gtex_readin("ADRNLG", "adrenal_gland")


comb_tissues <- function(tissue_list, target_table){
  #combine drugs that target the same gene to avoid duplicate gene targets
  target_table <- target_table %>% dplyr::group_by(Drug_Targets) %>% mutate(Drugs = stringr::str_flatten(Drug, ", "))
  #remove gene target duplicates
  target_table <- dplyr::distinct(target_table, Drug_Targets, .keep_all = TRUE)
  comb <- do.call("rbind", tissue_list)
  #filter for targets of prioritized drugs
  comb <- dplyr::left_join(target_table, comb, by = c("Drug_Targets" = "HUGO_gene_id"))
  
return(comb)
}

targetsintissues <- function(combinedtissues){
  #need wide format, genes by tissues
  tissues_exp <- combinedtissues %>% dplyr::select(Drug_Targets, Tissue, `MASH beta`) %>% tidyr::pivot_wider(names_from = Tissue, values_from = `MASH beta`)
  #remove gene target duplicates
  #tissues_exp <- dplyr::distinct(tissues_exp, Drug_Targets, .keep_all = TRUE)
  
  #make new table of just gene target sex-biased DE for each tissue
  #mat <- cbind(tissues_exp[,11:length(tissues_exp)])
  #remove NA column from NA's in original "tissue" column
  mat <- tissues_exp[,-2]
  mat <- tibble::column_to_rownames(mat, var = "Drug_Targets")

return(mat)
}
```

read in gtex sex-specific, tissue-specific data and add column of tissue group (top 500 sex-biased genes, pre-filtered by local false sign rate (LFSR) ≤ 0.05)    
looking at supplementary table 3, looks like > 0 = F bias, < 0 = M
```{r}
adrenal_gland <- gtex_readin("ADRNLG")

colon_transverse <- gtex_readin("CLNTRN")

heart_atrial_appendage <- gtex_readin("HRTAA")

heart_left_ventricle <- gtex_readin("HRTLV")

minor_salivary_gland <- gtex_readin("SLVRYG")

pituitary <- gtex_readin("PTTARY")

small_intestine <- gtex_readin("SNTTRM")

brain_cortex <- gtex_readin("BRNCTXA")

brain_frontal_cortex <- gtex_readin("BRNCTXB")

brain_hippocampus <- gtex_readin("BRNHPP")

brain_hypothalamus <- gtex_readin("BRNHPT")

brain_spinalcord <- gtex_readin("BRNSPC")

colon_sigmoid <- gtex_readin("CLNSGM")

kidney <- gtex_readin("KDNCTX")

liver <- gtex_readin("LIVER")

lung <- gtex_readin("LUNG")

muscle <- gtex_readin("MSCLSK")

pancreas <- gtex_readin("PNCREAS")

stomach <- gtex_readin("STMACH")

whole_blood <- gtex_readin("WHLBLD")

```

read in previously prioritized drugs
```{r}
top_16drugs <- read.csv("/Users/eramsey/Downloads/kidney_drugrepurposing/pkd_drugrepurposing/res/sigsearch_outputs/drug_summaries_oe_cystic.csv", row.names = 1)
#separate out drug targets
top_16drugs <- tidyr::separate_rows(top_16drugs, Drug_Targets, sep = "; ")
length(unique(top_16drugs$Drug_Targets))
```



comb_tissues: function to combine tissue tables and filter genes for drug targets  
targetsintissues:  function to data wrangle output of comb_tissues, returns a matrix of just MASH beta values (from sex bias LFC, < 0 = M, > 0 = F) for each tissue 
```{r}
tissue_list <- list(adrenal_gland, brain_cortex, heart_atrial_appendage, heart_left_ventricle, minor_salivary_gland, pituitary, small_intestine, brain_frontal_cortex, brain_hippocampus, brain_hypothalamus, brain_spinalcord, colon_sigmoid, colon_transverse, kidney, liver, lung, muscle, pancreas, stomach, whole_blood)

combined_tissues <- comb_tissues(tissue_list, top_16drugs)

target_tissue_exp <- targetsintissues(combined_tissues)

```


drug targets that are sex biased in at least 1 tissue
```{r}
target_tissue_exp <- filter(target_tissue_exp, rowSums(is.na(target_tissue_exp)) != ncol(target_tissue_exp))

```

table of na frequency of sex-biased de across tissues
```{r}
target_tissue_exp <- target_tissue_exp %>% tibble::rownames_to_column(var = "targets") %>%
  rowwise %>%
  mutate(NA_per_row = sum(is.na(across()))) %>% 
  #mutate(Median_Bias = median(across())) #%>% 
  arrange(NA_per_row) #%>% tibble::column_to_rownames(var = "targets")

```

drug-target annotations (color of simple annotation should be specified as a list with names for which names in the color list (here it is type in following example) correspond to the names in the data frame. Each color vector should better has names as well to map to the levels of annotations https://bioconductor.riken.jp/packages/3.8/bioc/vignettes/ComplexHeatmap/inst/doc/s4.heatmap_annotation.html )
```{r}
#remove gene dupicates and filter for genes with sex bias in at least 1 tissue
drug_categ <- combined_tissues %>% dplyr::select(Drug_Targets, Drugs) %>% dplyr::distinct( Drug_Targets, .keep_all = TRUE) %>% dplyr::filter(Drug_Targets %in% row.names(target_tissue_exp))
#make annotation table
#drug_categ <- as.data.frame(factor(drug_categ$Drugs), row.names = drug_categ$Drug_Targets)

#drug_categ <- as.data.frame(drug_categ$Drugs, row.names = drug_categ$Drug_Targets)
#drug_ha <- factor(drug_categ$Drugs, labels = drug_categ$Drug_Targets)
#names(drug_ha) <- drug_categ$Drug_Targets
#drug_categ <- as.data.frame(Drugs = drug_categ, row.names = combined_tissues$Drug_Targets)
#pdf(file = "colorpallette.pdf")
#scales::show_col(viridis(length(drug_categ$Drug_Targets), option = "D")) #selecting colors
#dev.off()

#named list ?
drug_ha <- drug_categ$Drugs
names(drug_ha) <- drug_categ$Drug_Targets
drug_ha <- rowAnnotation(Drugs = drug_ha)
#drug_ha <- rowAnnotation(Drugs = drug_ha, col = list(Drugs = colorRamp2(c(0, length(drug_categ$Drug_Targets)), c("white", "purple"))))

Drug_Targets <- drug_categ$Drug_Targets
names(Drug_Targets) <- Drug_Targets

```

colors for annotations
```{r}
#drugtarget_ha <- rowAnnotation(Drug_Targets = Drug_Targets, col = list(Drugs = c("ADRA1A" = "#440154", "ADRA2C" = "#470D60", "PRLR" = "#481A6CFF", "CYP3A5" = "#482576FF", "IRAK1"  = "#472F7DFF", "LTK" = "#443A83FF", "KCND2" = "#414487FF", "KCNH2" = "#3D4D8AFF", "SLC6A3" = "#39568CFF", "KCNA1" = "#35608DFF", "OPRD1" = "#31688EFF", "KIF1A" = "#2D718EFF", "KIF1C" = "#2A788EFF", "MAP4" = "#27808EFF", "CACNA2D1" = "#23888EFF", "CYP3A7" = "#21908CFF", "AOX1" = "#1F988BFF", "ABCD2" = "#1FA188FF", "ABCD3" = "#22A884FF", "CETP"  = "#29AF7FFF", "CRAT" = "#35B779FF", "LPL" = "#43BF71FF", "PPARG" = "#54C568FF", "SERPINE1" = "#66CB5DFF", "CHRNA1" = "#7AD151FF", "CHRNA5" = "#8FD744FF", "CYP2U1" = "#A5DB36FF", "FMO5" = "#BBDF27FF", "PNMT" = "#D2E21BFF", "SELP" = "#E9E519FF", "AR" = "#FDE725FF")))
drugtarget_ha <- rowAnnotation(Drug_Targets = Drug_Targets, col = list(Drugs = c("ADRA1A" = magma(31)[5], "ADRA2C" = magma(31)[7], "PRLR" = magma(31)[10], "CYP3A5" = magma(31)[12], "IRAK1"  = magma(31)[15],  "LTK" = magma(31)[18], "KCND2" = magma(31)[21], "KCNH2" = "#3D4D8AFF", "SLC6A3" = "#39568CFF", "KCNA1" = "#35608DFF", "OPRD1" = "#31688EFF", "KIF1A" = "#2D718EFF", "KIF1C" = "#2A788EFF", "MAP4" = "#27808EFF", "CACNA2D1" = "#23888EFF", "CYP3A7" = "#21908CFF", "AOX1" = "#1F988BFF", "ABCD2" = "#1FA188FF", "ABCD3" = "#22A884FF", "CETP"  = "#29AF7FFF", "CRAT" = "#35B779FF", "LPL" = "#43BF71FF", "PPARG" = "#54C568FF", "SERPINE1" = "#66CB5DFF", "CHRNA1" = "#7AD151FF", "CHRNA5" = "#8FD744FF", "CYP2U1" = "#A5DB36FF", "FMO5" = "#BBDF27FF", "PNMT" = "#D2E21BFF", "SELP" = "#E9E519FF", "AR" = "#FDE725FF")))


colorfill <- c("ADRA1A" = "#440154", "ADRA2C" = "#470D60", "PRLR" = "#481A6CFF", "CYP3A5" = "#482576FF", "IRAK1"  = "#472F7DFF", "LTK" = "#443A83FF", "KCND2" = "#414487FF", "KCNH2" = "#3D4D8AFF", "SLC6A3" = "#39568CFF", "KCNA1" = "#35608DFF", "OPRD1" = "#31688EFF", "KIF1A" = "#2D718EFF", "KIF1C" = "#2A788EFF", "MAP4" = "#27808EFF", "CACNA2D1" = "#23888EFF", "CYP3A7" = "#21908CFF", "AOX1" = "#1F988BFF", "ABCD2" = "#1FA188FF", "ABCD3" = "#22A884FF", "CETP"  = "#29AF7FFF", "CRAT" = "#35B779FF", "LPL" = "#43BF71FF", "PPARG" = "#54C568FF", "SERPINE1" = "#66CB5DFF", "CHRNA1" = "#7AD151FF", "CHRNA5" = "#8FD744FF", "CYP2U1" = "#A5DB36FF", "FMO5" = "#BBDF27FF", "PNMT" = "#D2E21BFF", "SELP" = "#E9E519FF", "AR" = "#FDE725FF")
#tissue_ha = HeatmapAnnotation(df = data.frame(tissue = colnames(target_tissue_exp), group = ifelse(target_tissue_exp starts"")), col = list(age = colorRamp2(c(0, 20), c("white", "red"))))
#df = data.frame(tissue = colnames(target_tissue_exp), group = ifelse(target_tissue_exp starts_with()))
#my_color = list(Direction = c(aki_down = "#575C6DFF", aki_up = "#FFEA46FF"), Cystic_Status = c(Precystic = "#00204DFF", Cystic = "#A69D75FF"))
my_color = list(Target = c("ADRA1A" = "#440154", "ADRA2C" = "#470D60", "PRLR" = "#481A6CFF", "CYP3A5" = "#482576FF", "IRAK1"  = "#472F7DFF", "LTK" = "#443A83FF", "KCND2" = "#414487FF", "KCNH2" = "#3D4D8AFF", "SLC6A3" = "#39568CFF", "KCNA1" = "#35608DFF", "OPRD1" = "#31688EFF", "KIF1A" = "#2D718EFF", "KIF1C" = "#2A788EFF", "MAP4" = "#27808EFF", "CACNA2D1" = "#23888EFF", "CYP3A7" = "#21908CFF", "AOX1" = "#1F988BFF", "ABCD2" = "#1FA188FF", "ABCD3" = "#22A884FF", "CETP"  = "#29AF7FFF", "CRAT" = "#35B779FF", "LPL" = "#43BF71FF", "PPARG" = "#54C568FF", "SERPINE1" = "#66CB5DFF", "CHRNA1" = "#7AD151FF", "CHRNA5" = "#8FD744FF", "CYP2U1" = "#A5DB36FF", "FMO5" = "#BBDF27FF", "PNMT" = "#D2E21BFF", "SELP" = "#E9E519FF", "AR" = "#FDE725FF"))

```


data wrangling for alluvial plot
```{r}
#merge sex bias values from each tissue with drug table?
#top_16drugs_sb <- merge(top_16drugs, )

#top_16drugs_sb <- dplyr::filter(top_16drugs, Drug_Targets %in% target_tissue_exp$targets) #row.names(target_tissue_exp))
top_16drugs_sb <- merge(top_16drugs, target_tissue_exp, by.x = "Drug_Targets", by.y = "targets")
#top_16drugs_sb <- mutate(top_16drugs_sb, Drug_Targets = forcats::fct_reorder(Drug_Targets, NA_per_row))
top_16drugs_sb <- top_16drugs_sb %>% mutate(Drug_Targets = forcats::fct_reorder(Drug_Targets, NA_per_row)) 
#top_16drugs_sb <- dplyr::inner_join(target_tissue_exp, top_16drugs, by = c("targets" = "Drug_Targets"))
```

plot alluvial plot
```{r}
target_alluvial <- top_16drugs_sb %>% 
  #arrange(NA_per_row) %>%
  mutate(Drug_Targets = forcats::fct_reorder(Drug_Targets, NA_per_row)) %>%
  #ggplot(aes(x = NA_per_row, stratum = targets, alluvium = Drug)) +
  ggplot(aes( axis1 = Drug_Targets, axis2 = Drug)) +
  geom_alluvium(color = "black", aes(fill = Drug_Targets), show.legend=FALSE) + # , colour = trend)) + 
  geom_stratum(width = 3/12, color = "black") + #the border of the stratum
  scale_fill_viridis_d(aesthetics = c("colour", "fill")) +
  #scale_fill_manual(values = colorfill) +
  scale_x_discrete(limits = c("Drug_Targets", "Drug"),
                   expand = c(0.05, .05)) +
  geom_text(stat = "stratum", size = 4.5, fontface = "bold", aes(label = after_stat(stratum))) +
  theme(axis.title.x = element_blank(), plot.title = element_text(size = 16, face = "bold", hjust = 0.5, vjust = -5), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank()) + ggtitle(label = "Drugs Targetting Sex-Biased Genes") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank() )

target_alluvial

ggsave("/Users/eramsey/Downloads/PKD/kidney_drugrepurposing/figures/gtex_sexbiasDEG_drugtarget_alluvial.png", target_alluvial, width = 6, height = 10)
```

heatmap data wrangling
```{r}
#target_tissue_exp_hp <- target_tissue_exp %>% mutate(targets = forcats::fct_reorder(targets, NA_per_row)) %>% tibble::column_to_rownames(var = "targets") %>% dplyr::select(!"NA_per_row") 
target_tissue_exp_hp <- top_16drugs_sb %>% dplyr::distinct(Drug_Targets, .keep_all = TRUE)  %>% arrange(NA_per_row) %>% tibble::column_to_rownames(var = "Drug_Targets") %>% dplyr::select(c(sort(na.omit(combined_tissues$Tissue)))) 

tissue_annot <- data.frame(Tissue = na.omit(combined_tissues$Tissue), Group = na.omit(combined_tissues$`Tissue group`)) 
#tissue_annot <- complete.cases(tissue_annot)
#tissue_annot <- dplyr::filter(tissue_annot, Tissue %in% colnames(target_tissue_exp_hp))
tissue_annot <- tissue_annot %>% dplyr::distinct(Tissue, .keep_all = TRUE) %>% dplyr::arrange(Tissue) %>% tibble::column_to_rownames("Tissue")

#color palette for annotations
#color_anno <- viridis(n= length(row.names(tissue_annot)), alpha = 1, begin = 0, end = 1, option = "viridis")
#names(color_anno) <- row.names(tissue_annot)

color_anno <- list(Tissue = row.names(tissue_annot), Group = c(viridis(n= length(unique(tissue_annot$Group)), alpha = 1, begin = 0, end = 1, option = "viridis")))
#tissue_annot$Color =  c(viridis(n= length(unique(tissue_annot$Group))))

#[17] "#94D840FF" "#B8DE29FF" "#DCE318FF" "#FDE725FF"
color_anno <- list(Group = c("Adrenal Gland" = "#440154FF" , "Brain" = "#481D6FFF", "Colon" = "#453581FF", "Heart" = "#3D4D8AFF", "Kidney" = "#34618DFF", "Liver" = "#2B748EFF", "Lung" = "#24878EFF", "Salivary Gland" = "#1F998AFF","Muscle" = "#25AC82FF", "Pancreas" = "#40BC72FF", "Pituitary" = "#67CC5CFF", "Small Intestine" = "#97D83FFF", "Stomach" = "#CBE11EFF", "Blood" = "#FDE725FF"))

```
 

color breaks (https://gist.github.com/xdoan/e39ded4d4680354f063ff16a3318e1e5)
```{r}
myBreaks <- c(
              seq(min(target_tissue_exp_hp, na.rm=T), 0, length.out=ceiling(3/3) + 1), #minimum to 0 log2FC
              seq((-1*min(target_tissue_exp_hp, na.rm=T))/3, (-1*min(target_tissue_exp_hp, na.rm=T)), length.out=floor(3/3)), #0 to absolute value of minimum
              seq((-1*min(target_tissue_exp_hp, na.rm=T))+max(target_tissue_exp_hp, na.rm=T)/3, 
                  max(target_tissue_exp_hp, na.rm=T), length.out=floor(3/3)) # absolute value of min to max
              )

paletteLength <- 100

myColor <- colorRampPalette(c("darkgreen", "white", "blueviolet"))(paletteLength)

myBreaks <- c(seq(min(target_tissue_exp_hp, na.rm=T), 0, length.out=ceiling(paletteLength/2) + 1), 
              seq(max(target_tissue_exp_hp, na.rm=T)/paletteLength, max(target_tissue_exp_hp, na.rm=T), length.out=floor(paletteLength/2)))



```

plot heatmap
```{r}
png("/Users/eramsey/Downloads/PKD/kidney_drugrepurposing/figures/gtex_sexbiasDEG_drugtarget_heatmap.png", width = 25, height = 40, units = "cm", res = 300)

ComplexHeatmap::pheatmap(target_tissue_exp_hp, annotation_colors = color_anno, #annotation_row = Target, 
                         #annotation_col = colnames(target_tissue_exp), 
                         cluster_rows = FALSE, cluster_cols = FALSE, show_rownames = TRUE, labels_col = stringr::str_trunc(colnames(target_tissue_exp_hp), 18), color = myColor, #color = brewer.pal(100, name = "PRGn"), 
                         fontsize = 20, fontface = "bold", fontsize_row = 16, fontsize_col = 16, annotation_col = tissue_annot,# viridis(n= 10000, alpha = 1, begin = 0, end = 1, option = "viridis"), 
                         border_color = "NA", main = "Sex-Biased DE of \n Drug Targets by Tissue", angle_col = "90", breaks = myBreaks, heatmap_legend_param = list(title = "Expression", at = c(-1, 0, 1), labels = c("Male", "None", "Female")))
dev.off()
```


#### Versions
```{r}
R.Version()
```


```{r}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```
  
